### 2012-03-30 mc/plot-insample-size.Rnw
Collapsed hold history into a single commit

\documentclass{article}
\SweaveOpts{prefix.string=plots/ftest}
\usepackage[margin=1in]{geometry}

\begin{document}
Basic file that generates plots for size of the F-test

<<fig=TRUE>>=
library(fwPackage)
library(lattice)

dbc <- dbConnect(dbDriver("SQLite"), dbname = "simulations.db")

print(dotplot(norm ~ reject | interaction(altlabel,nlabel, sep = ", "),
              dat = dbGetQuery(dbc, "select reject, 'n=' || n as nlabel, altlabel, norm from viewftest 
                                     where label='size' and norm in (0,1)"),
              xlab = "Rejection Probability",
              xlim = c(-.1, 1.1),
              panel = function(x, y, ...) {
                panel.dotplot(x, y, ..., type = "n")
                panel.text(x, as.numeric(y), labels = sprintf("%.2f", x))
              },
              par.settings = standard.theme(color=FALSE),
              par.strip.text = list(cex = .65),
              layout = c(2,4),
              index.cond = list(c(3,4,7,8,5,6,1,2)),
              main = "Rejection probability for F-test at 10% (size)\nbenchmark has smaller generalization error"))
@

<<>>=
dbDisconnect(dbc)
@
\end{document}