### 2012-03-30 mc/plot-interval.Rnw

\documentclass{article}
\SweaveOpts{prefix.string=plots/interval}
\usepackage[margin=1in]{geometry}

\begin{document}
Basic file that generates plots for the out-of-sample $t$-test
centered at $E_{R}\bar D_{1}$

<<>>=
library(fwPackage)
library(lattice)

dbc <- dbConnect(dbDriver("SQLite"), dbname = "simulations.db")
d <- dbGetQuery(dbc, "
select * from (select ntest, isim, idgp, transform, label, avg(reject) as reject
from interval 
where label in ('test.error.1', 'test.error.2', 'gen.error.1', 'gen.error.2')
and simIndex <= 500
group by ntest, isim, idgp, transform, label) s 
join nobs n join coefficients c on n.i=s.isim and c.i=s.idgp")
d$nlabel <- sprintf("n=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)

doxyplot <- function(dframe, label, maintext, ...) {
  dplot <- dframe[dframe$label==label,]
  print(xyplot(I(1 - reject) ~ ntest | interaction(nlabel, altlabel, normlabel, sep = ", "),
               data = dplot,
               ylab = "coverage", main = maintext,
##               ylim = c(-0.05,1.05),
               panel = function(x,y,...) {
                 panel.xyplot(x,y,...,type = "l")
                 panel.lines(c(1,600),c(.9,.9), col = rgb(0,0,0,0.5))
               },
               par.strip.text = list(cex = .65),
               index.cond = list(c(13,15,16,14,9,11,12,10,5,7,8,6,1,3,4,2)),
               par.settings = standard.theme(color=FALSE),...))
}
@

<<fig=TRUE>>=
doxyplot(d, "test.error.1", "Empirical coverage of one-sided interval")
@

<<fig=TRUE>>=
doxyplot(d, "test.error.2", "Empirical coverage of two-sided interval")
@

<<fig=TRUE>>=
doxyplot(d, "gen.error.1", "Empirical coverage of one-sided interval")
@

<<fig=TRUE>>=
doxyplot(d, "gen.error.2", "Empirical coverage of two-sided interval")
@

<<fig=TRUE>>=
doxyplot(d, "gen.error.2", "Details of empirical coverage of two-sided interval", 
         xlim = c(0,350), ylim = c(.75, .95))
@ 
<<>>=
dbDisconnect(dbc)
@
\end{document}