### 2012-03-30 mc/plot-oos-size.Rnw

\documentclass{article}
\SweaveOpts{prefix.string=plots/size}
\usepackage[margin=1in]{geometry}

\begin{document}
Basic file that generates plots for the out-of-sample $t$-test
centered at $E_{R}\bar D_{1}$

<<>>=
library(fwPackage)
library(lattice)

dbc <- dbConnect(dbDriver("SQLite"), dbname = "simulations.db")

d <- dbGetQuery(dbc, "
select * from (select ntest, isim, idgp, transform, label, avg(reject) as reject
from interval 
where label='size' and idgp in (1,2) and transform in ('mccracken', 'clarkwest', 'difference')
group by ntest, isim, idgp, transform, label) s 
join nobs n join coefficients c on n.i=s.isim and c.i=s.idgp")
d$nlabel <- sprintf("n=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)

dpower <- dbGetQuery(dbc, "
select * from (select ntest, isim, idgp, transform, label, avg(reject) as reject
from interval 
where label='power' and idgp in (2,3) and transform='difference'
group by ntest, isim, idgp, transform, label) s 
join nobs n join coefficients c on n.i=s.isim and c.i=s.idgp")
dpower$nlabel <- sprintf("n=%d", dpower$n)
dpower$normlabel <- sprintf("c=%d", dpower$norm)

# dboth <- rbind(d, dpower)

doxyplot <- function(dframe, maintext, ...) {
  print(xyplot(reject ~ ntest | interaction(nlabel,altlabel,normlabel, sep = ", "),
               data = dframe,
               ylab = "size", main = maintext,
               panel = function(x,y,...) {
                 panel.xyplot(x,y,...,type = "l")
                 panel.lines(c(1,600),c(.1,.1), col = rgb(0,0,0,0.5))
               },
               layout = c(4,4),
               par.strip.text = list(cex = .65),
               index.cond = list(c(13,15,16,14,9,11,12,10,5,7,8,6,1,3,4,2)),
               par.settings = standard.theme(color=FALSE)))
}
@

<<fig=TRUE>>=
doxyplot(d[d$transform == "clarkwest",], 
         "Empirical size of Clark and West test at 90% confidence")
@

<<fig=TRUE>>=
doxyplot(d[d$altlabel=="K=3" & d$transform == "mccracken",], 
         "K=3", "Empirical size of McCracken test at 90% confidence")
@

<<fig=TRUE>>=
doxyplot(d[d$transform == "difference",], 
         "K=3", "Empirical size of Naive OOS test at 90% confidence", auto.key = list(lines=TRUE, points=FALSE))
@

<<fig=TRUE>>=
doxyplot(dpower[dpower$transform == "difference",], 
         "Empirical power of Naive oos test at 90% confidence", auto.key = list(lines=TRUE, points=FALSE))
@

<<>>=
dbDisconnect(dbc)
@
\end{document}