### 2012-03-30 mc/plot-insample-size.Rnw
Typed up changes to paper

\documentclass{article}
\SweaveOpts{prefix.string=plots/ftest}
\usepackage[margin=1in]{geometry}
\usepackage{tikz}

\begin{document}
Basic file that generates plots for size of the F-test

<<fig=TRUE,width=5.5, height=4, eps=F, pdf=T>>=
library(fwPackage)
library(pgfSweave)
library(lattice)
ltheme <- canonical.theme(color = FALSE)     ## in-built B&W theme  
ltheme$strip.background$col <- "transparent" ## change strip bg  
lattice.options(default.theme = ltheme)      ## set as default

dbc <- dbConnect(dbDriver("SQLite"), dbname = "simulations.db")
d <- dbGetQuery(dbc, "select reject, 'T=' || n as nlabel, altlabel, norm from viewftest 
                                     where label='size' and norm in (0,1)")
d$norm <- as.factor(d$norm)

print(dotplot(norm ~ reject | interaction(altlabel,nlabel, sep = ", "),
              dat = d, las = 1,
              xlab = "Rejection Probability", ylab = "c",
              xlim = c(-.1, 1.1),
              panel = function(x, y, ...) {
                panel.dotplot(x, y, ..., type = "n")
                panel.text(x, as.numeric(y), labels = sprintf("%.2f", x))
              },
              par.strip.text = list(cex = .55),
              layout = c(2,4),
              index.cond = list(c(3,4,7,8,5,6,1,2))))
@

<<>>=
dbDisconnect(dbc)
@
\end{document}