### 2014-03-19 appendix.tex
Add recursive window simulations for separate appendix

\documentclass[11pt]{article}
\input{setup}

\title{Supplamental Appendix for Out-of-Sample Comparisons of Overfit
  Models}
\author{Gray Calhoun\thanks{email: gcalhoun@iastate.edu} \\
  Iowa State University}

\begin{document}
\maketitle

Fill in the text...

\begin{figure}
  \centering {\large Simulated Coverage of Recursive One-Sided DMW Interval for
    $\E_T\oosB$} \input{floats/mc-interval-generror1-rec.tex}
\caption{Simulated coverage of $\E_T \oosB$ at 90\% confidence using a
  one-sided interval based on the DMW
  OOS, plotted as a function of the fraction of
  observations used in the test sample, $P/T$.  The solid horizontal
  line denotes the intervals' nominal coverage.}
\label{fig:interval-T}
\end{figure}
\clearpage

\begin{figure}
  \centering {\large Simulated Rejection Probability of Recursive DMW Test Under
    $\E_T \oosB \leq 0$} \input{floats/mc-dmwsize-rec.tex}
  \caption{Simulated rejection probabilities for the DMW
    OOS $t$-test given $\E_T \oosB \leq 0$ with nominal
    size 10\%.  Values greater than 10\% indicate that the test
    rejects the benchmark model too often.  The solid horizontal line
    indicates the nominal rejection probability.  See
    Section~\ref{sec:simulation-design} for a discussion of the
    simulation design.}
  \label{fig:ttest-size}
\end{figure}

\begin{figure}
  \centering {\large Simulated Rejection Probability of Recursive Clark and West
    (2006, 2007) \\ Test Under $\E_T \oosB \leq 0$}
  \input{floats/mc-clarkwestsize-rec.tex}
  \caption{Simulated rejection probabilities for Clark and West's
    (2006, 2007) OOS test statistic given $\E_T \oosB
    \leq 0$ with nominal size 10\%.  Values greater than 10\% indicate
    that the test rejects the benchmark model too often.  The solid
    horizontal line indicates the nominal rejection probability.  See
    Section~\ref{sec:simulation-design} for a discussion of the
    simulation design.}
   \label{fig:clarkwest}
\end{figure}

\begin{figure}
  \centering {\large Simulated Rejection Probability of Recursive McCracken
    (2007) Test \\ Under $\E_T \oosB \leq 0$}
  \input{floats/mc-mccrackensize-rec.tex}
  \caption{Simulated rejection probabilities for McCracken's (2007)
    OOS test statistic given $\E_T \oosB \leq 0$ with
    nominal size 10\%.  Values greater than 10\% indicate that the
    test rejects the benchmark model too often.  The solid horizontal
    line indicates the nominal rejection probability.  See
    Section~\ref{sec:simulation-design} for a discussion of the
    simulation design.}
  \label{fig:mccracken}
\end{figure}

\begin{figure}
  \centering {\large Simulated Rejection Probability of Recursive DMW Test Under
    $\E_T \oosB > 0$} \input{floats/mc-dmwpower-rec.tex}
  \caption{Simulated rejection probabilities for the DMW
    OOS $t$-test given $\E_T \oosB > 0$ with nominal
    size 10\%.  Values greater than 10\% indicate that the test
    rejects the benchmark model too often.  The solid horizontal line
    indicates the nominal rejection probability.  See
    Section~\ref{sec:simulation-design} for a discussion of the
    simulation design.}
  \label{fig:ttest-power}
\end{figure}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

% LocalWords:  McCracken's kilian anatolyev rossi REStud JAppliedEconometrics
% LocalWords:  goyal welch rfs reputational calhoun Berbee's mixingales Rt tS
% LocalWords:  Heteroskedasticity Autocorrelation Tt iT th berbee iS de dx ixz
% LocalWords:  DMW mcleish Rs eq ftestreject lightgray veclen RSQLite zeileis
% LocalWords:  heteroskedasticity hothorn RNews JStatSoftware harrell campbell
% LocalWords:  thompson bivariate Welch's multicollinearity Newey newey tw vw
% LocalWords:  merlevede cltvar dR Scwarz davidson indices lrrrr waldtest Df jt
% LocalWords:  dmw McCracken MSE OLS mse inoue huber elemstatlearning efron sw
% LocalWords:  anova akritas jong normalpart coefpart Jong's rproject rsqlite
% LocalWords:  unrestrictive datasets Heyde's ib oos Mcc MeR StW InK DiM CCS lm
% LocalWords:  ClM CoS ClW GiW GiR GoW Cla HTF Efr BoB AkA AkP DeD Mcl Dej Rde
% LocalWords:  Sar Zeh Zei Har CaT NeW MeP HaH AllRefs Diebold jel Whi RoW brc
% LocalWords:  overrejection overreject fpe JoD iR PeT Giacomini WN Wishart T'V
% LocalWords:  premultiply const overrejects overrejecting BWB Gia Schwarz HHK
% LocalWords:  Haf Anatolyev's nonstochastic T'X Sweave tikzDevice pgfSweave