### 2014-03-25 floats/empirics-insample-tuned.tex

\renewcommand\insampletests\relax

\renewcommand\insamplecoefs{%
  \begin{table}[tb]
    \begin{tabularx}{\textwidth}{lCCCC}
      \toprule
      & \multicolumn{2}{c}{Coefficient estimates}
      & \multicolumn{2}{c}{$p$-values} \\
      \cmidrule(lr){2-3} \cmidrule(l){4-5}
      & Full & Univariate & Full & Univariate \\
      \midrule Stock market variance & $\enskip4.61$ & $\enskip0.36$ & $0.081$ & $0.870$ \\
      Book to market ratio & $\enskip4.46$ & $\enskip4.81$ & $0.251$ & $0.027$ \\
      Dividend to price ratio (log) & $\enskip4.40$ & $\enskip3.91$ & $0.400$ & $0.073$ \\
      Long term rate & $\enskip3.78$ & $\enskip2.91$ & $0.035$ & $0.184$ \\
      Long term yield & $\enskip3.23$ & $\!\!-1.25$ & $0.438$ & $0.569$ \\
      Earnings to price ratio (log) & $\enskip0.79$ & $\enskip3.50$ & $0.849$ & $0.109$ \\
      Inflation rate & $\enskip0.63$ & $\enskip0.69$ & $0.696$ & $0.754$ \\
      Default return spread & $\enskip0.53$ & $\!\!-1.74$ & $0.795$ & $0.429$ \\
      Net equity expansion & $\!\!-1.88$ & $\!\!-4.10$ & $0.626$ & $0.060$ \\
      Treasury Bill rate & $\!\!-4.38$ & $\!\!-2.47$ & $0.316$ & $0.260$ \\
      Percent equity issuing & $\!\!-5.56$ & $\!\!-5.10$ & $0.040$ & $0.019$ \\
      Default yield spread & $\!\!-7.79$ & $\enskip0.79$ & $0.042$ & $0.720$ \\
      \\
      Robust $F$-statistic ($12$/$69$ df) & \enskip $3.37$ & & $0.001$ \\
      $R^2$ & $\enskip 0.23$ \\
      Adjusted $R^2$ & $\enskip 0.10$ \\
      \bottomrule \end{tabularx}
    \caption{Coefficient estimates and model fit for the univariate and full
      models described in Section \ref{sec:empirics}. The ``Full'' column
      lists the coefficient estimates from regressing the equity premium on
      all of the variables listed; the ``Univariate'' column lists the
      coefficient estimates from a univariate regression of the equity premium
      on the variable alone. All of the regressions include a constant, but
      its estimate is not listed. The equity premium is expressed in basis
      points and all of the regressors are studentized.
      These models are estimated using U.S. annual data from 1928--2009
      (81 observations). The standard errors and robust $F$-statistic
      are calculated using a Newey-West HAC estimator with two lags,
      implemented in the \emph{sandwich} and \emph{lmtest} R packages
      \protect\citep{NeW:87,ZeH:02,Zei:04}.}
    \label{tab:gwinsample}
  \end{table}
}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End: