### 2014-03-27 appendix.tex
Correct stupid spelling mistake in title of appendix

\documentclass[12pt]{article}
\input{setup}

\title{Supplemental Appendix for ``Out-of-Sample Comparisons of
  Overfit Models''}
\author{Gray Calhoun\thanks{email: \protect\url{gcalhoun@iastate.edu}} \\
  Iowa State University}

\begin{document}
\maketitle

This appendix presents simulation results that accompany
\citet{Cal:14}. The Monte Carlo here is exactly the same as presented
in \citet{Cal:14}, except that a recursive window is used. Please read
the original paper for further details.

\begin{itemize}

\item Figure 1 in this Appendix corresponds to Figure 3 in the
  original paper.
\item Figure 2 corresponds to Figure 4 in the paper.
\item Figure 3 corresponds to Figure 5 in the paper.
\item Figure 4 corresponds to Figure 6 in the paper.
\item Figure 5 corresponds to Figure 7 in the paper.

\end{itemize}

\nocite{DiM:95,Wes:96,Mcc:07,ClW:06,ClW:07}
\bibliographystyle{abbrvnat}
\bibliography{references}

\begin{figure}\centering
  {\large Coverage of recursive DMW OOS interval in simulations}
  \input{floats/mc-interval-generror1-rec.tex}
  \caption{Simulated coverage of $\E_T \oosB$ at 90\% confidence using
    a one-sided interval based on the recursive DMW \oost\ test,
    plotted as a function of the fraction of observations used in the
    test sample, $P/T$.  The solid horizontal line denotes the
    intervals' nominal coverage.}
  \label{fig:interval-T}
\end{figure}

\begin{figure}\centering
  {\large Size of recursive DMW \oost\ test in simulations}
  \input{floats/mc-dmwsize-rec.tex}
  \caption{Simulated rejection probabilities for the recursive DMW
    \oost\ test under the null hypothesis that the benchmark model is
    more accurate, $\E_T \oosB \leq 0$. Nominal size is 10\% and is
    marked with a solid horizontal line. Values greater than 10\%
    indicate that the test rejects the benchmark model too often.}
  \label{fig:ttest-size}
\end{figure}

\begin{figure}\centering
  {\large Size of recursive Clark-West OOS test in simulations}
  \input{floats/mc-clarkwestsize-rec.tex}
  \caption{Simulated rejection probabilities for Clark and West's
    (2006, 2007) recursive OOS test statistic under the null
    hypothesis that the benchmark model is more accurate, $\E_T \oosB
    \leq 0$. Nominal size is 10\% and is marked with a solid
    horizontal line. Values greater than 10\% indicate that the test
    rejects the benchmark model too often.}
  \label{fig:clarkwest}
\end{figure}

\begin{figure}\centering
  {\large Size of recursive McCracken \oost\ test in simulations}
  \input{floats/mc-mccrackensize-rec.tex}
  \caption{Simulated rejection probabilities for McCracken's (2007)
    recursive \oost\ test under the null hypothesis that the benchmark
    model is more accurate, $\E_T \oosB \leq 0$. Nominal size is 10\%
    and is marked with a solid horizontal line. Values greater than
    10\% indicate that the test rejects the benchmark model too
    often.}
  \label{fig:mccracken}
\end{figure}

\begin{figure}\centering
  {\large Power of recursive DMW \oost\ test in simulations}
  \input{floats/mc-dmwpower-rec.tex}
  \caption{Simulated rejection probabilities for the recursive DMW
    \oost\ test under the alternative that the benchmark model is less
    accurate, $\E_T \oosB > 0$. Nominal size is 10\% and is marked
    with a solid horizontal line. Values greater than 10\% indicate
    that the test rejects the benchmark model too often.}
  \label{fig:ttest-power}
\end{figure}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "appendix"
%%% End: