### 2016-03-09 appendix.tex
Include appendix with main paper for convenience to referees

\documentclass[12pt,draft]{article}
\input{setup}

\title{Supplemental Appendix for ``Out-of-Sample Comparisons of
  Overfit Models''}
\author{Gray Calhoun\thanks{email: \protect\url{gcalhoun@iastate.edu}} \\
  Iowa State University}

\date{\VERSION}

\begin{document}
\maketitle

\input{appendix_contents.tex}

\bibliographystyle{abbrvnat}
\bibliography{latex-tools-0.2.1/references}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "appendix.tex"
%%% TeX-command-extra-options: "-shell-escape"
%%% End: