### 2017-04-21 floats/empirics-oos-ind-ks.tex
Add TeX files with simulation and empirical results to repo
I've decided to add _a few_ generated files to the main repository to
make my life easier; namely any graphics or tables that are included
in the main paper.

% Created by tikzDevice version 0.6.2-92-0ad2792 on 2014-03-27 04:41:27
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (397.48,144.54);
\begin{scope}
\path[clip] ( 50.40, 61.20) rectangle (397.48,131.34);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] ( 63.25,128.74) --
	( 69.44,113.58) --
	( 75.62,116.70) --
	( 81.80,101.94) --
	( 87.98,102.92) --
	( 94.16,109.79) --
	(100.34,103.33) --
	(106.52,117.62) --
	(112.70,122.31) --
	(118.88,123.05) --
	(125.06,107.38) --
	(131.24,103.26) --
	(137.42,108.90) --
	(143.60, 81.56) --
	(149.78, 73.07) --
	(155.96, 70.30) --
	(162.14, 68.61) --
	(168.32, 68.52) --
	(174.50, 68.01) --
	(180.68, 66.99) --
	(186.86, 67.19) --
	(193.04, 68.04) --
	(199.22, 65.70) --
	(205.40, 65.82) --
	(211.58, 65.23) --
	(217.76, 65.25) --
	(223.94, 65.20) --
	(230.12, 65.91) --
	(236.30, 65.56) --
	(242.48, 65.28) --
	(248.66, 65.84) --
	(254.84, 65.68) --
	(261.02, 65.76) --
	(267.20, 65.83) --
	(273.38, 65.88) --
	(279.57, 64.78) --
	(285.75, 64.82) --
	(291.93, 64.59) --
	(298.11, 64.52) --
	(304.29, 64.56) --
	(310.47, 64.63) --
	(316.65, 64.48) --
	(322.83, 64.47) --
	(329.01, 64.52) --
	(335.19, 64.47) --
	(341.37, 64.51) --
	(347.55, 64.52) --
	(353.73, 64.57) --
	(359.91, 64.47) --
	(366.09, 64.51) --
	(372.27, 64.63) --
	(378.45, 64.81) --
	(384.63, 65.01);

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] ( 63.25, 63.80) --
	( 69.44, 63.80) --
	( 75.62, 63.80) --
	( 81.80, 63.80) --
	( 87.98, 63.80) --
	( 94.16, 63.80) --
	(100.34, 63.80) --
	(106.52, 63.80) --
	(112.70, 63.80) --
	(118.88, 63.80) --
	(125.06, 63.80) --
	(131.24, 63.80) --
	(137.42, 63.80) --
	(143.60, 63.80) --
	(149.78, 63.80) --
	(155.96, 63.80) --
	(162.14, 63.80) --
	(168.32, 63.80) --
	(174.50, 63.80) --
	(180.68, 63.80) --
	(186.86, 63.80) --
	(193.04, 63.80) --
	(199.22, 63.80) --
	(205.40, 63.80) --
	(211.58, 63.80) --
	(217.76, 63.80) --
	(223.94, 63.80) --
	(230.12, 63.80) --
	(236.30, 63.80) --
	(242.48, 63.80) --
	(248.66, 63.80) --
	(254.84, 63.80) --
	(261.02, 63.80) --
	(267.20, 63.80) --
	(273.38, 63.80) --
	(279.57, 63.80) --
	(285.75, 63.80) --
	(291.93, 63.80) --
	(298.11, 63.80) --
	(304.29, 63.80) --
	(310.47, 63.80) --
	(316.65, 63.80) --
	(322.83, 63.80) --
	(329.01, 63.80) --
	(335.19, 63.80) --
	(341.37, 63.80) --
	(347.55, 63.80) --
	(353.73, 63.80) --
	(359.91, 63.80) --
	(366.09, 63.80) --
	(372.27, 63.80) --
	(378.45, 63.80) --
	(384.63, 63.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 12.00, 96.27) {Full model};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25, 61.20) -- (372.27, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25, 61.20) -- ( 63.25, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.06, 61.20) -- (125.06, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.86, 61.20) -- (186.86, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.66, 61.20) -- (248.66, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.47, 61.20) -- (310.47, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.27, 61.20) -- (372.27, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 63.25, 39.60) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (125.06, 39.60) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (186.86, 39.60) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (248.66, 39.60) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (310.47, 39.60) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (372.27, 39.60) {70};
\end{scope}
\begin{scope}
\path[clip] ( 50.40, 61.20) rectangle (397.48,131.34);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25,128.74) --
	( 69.44,113.58) --
	( 75.62,116.70) --
	( 81.80,101.94) --
	( 87.98,102.92) --
	( 94.16,109.79) --
	(100.34,103.33) --
	(106.52,117.62) --
	(112.70,122.31) --
	(118.88,123.05) --
	(125.06,107.38) --
	(131.24,103.26) --
	(137.42,108.90) --
	(143.60, 81.56) --
	(149.78, 73.07) --
	(155.96, 70.30) --
	(162.14, 68.61) --
	(168.32, 68.52) --
	(174.50, 68.01) --
	(180.68, 66.99) --
	(186.86, 67.19) --
	(193.04, 68.04) --
	(199.22, 65.70) --
	(205.40, 65.82) --
	(211.58, 65.23) --
	(217.76, 65.25) --
	(223.94, 65.20) --
	(230.12, 65.91) --
	(236.30, 65.56) --
	(242.48, 65.28) --
	(248.66, 65.84) --
	(254.84, 65.68) --
	(261.02, 65.76) --
	(267.20, 65.83) --
	(273.38, 65.88) --
	(279.57, 64.78) --
	(285.75, 64.82) --
	(291.93, 64.59) --
	(298.11, 64.52) --
	(304.29, 64.56) --
	(310.47, 64.63) --
	(316.65, 64.48) --
	(322.83, 64.47) --
	(329.01, 64.52) --
	(335.19, 64.47) --
	(341.37, 64.51) --
	(347.55, 64.52) --
	(353.73, 64.57) --
	(359.91, 64.47) --
	(366.09, 64.51) --
	(372.27, 64.63) --
	(378.45, 64.81) --
	(384.63, 65.01);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 63.80) -- ( 50.40,119.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 63.80) -- ( 44.40, 63.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 77.75) -- ( 44.40, 77.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 91.71) -- ( 44.40, 91.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40,105.66) -- ( 44.40,105.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40,119.62) -- ( 44.40,119.62);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40, 61.04) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40, 75.00) {1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40, 88.95) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40,102.91) {3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40,116.86) {4};
\end{scope}
\end{tikzpicture}