### 2017-04-21 floats/empirics-oos-ind-pm.tex

% Created by tikzDevice version 0.6.2-92-0ad2792 on 2014-03-27 04:41:27
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (397.48,144.54);
\begin{scope}
\path[clip] ( 50.40, 61.20) rectangle (397.48,131.34);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] ( 63.25, 97.34) --
	( 69.44, 97.88) --
	( 75.62, 98.01) --
	( 81.80, 97.62) --
	( 87.98, 97.87) --
	( 94.16, 98.41) --
	(100.34, 98.76) --
	(106.52, 97.50) --
	(112.70, 97.95) --
	(118.88, 98.43) --
	(125.06, 97.50) --
	(131.24, 97.26) --
	(137.42, 97.93) --
	(143.60, 98.08) --
	(149.78, 98.74) --
	(155.96, 97.99) --
	(162.14, 98.74) --
	(168.32, 99.59) --
	(174.50,100.40) --
	(180.68, 99.28) --
	(186.86,100.08) --
	(193.04,100.89) --
	(199.22, 99.90) --
	(205.40,100.40) --
	(211.58,101.38) --
	(217.76,102.37) --
	(223.94,100.30) --
	(230.12, 94.81) --
	(236.30, 94.33) --
	(242.48, 94.91) --
	(248.66, 94.58) --
	(254.84, 95.42) --
	(261.02, 96.47) --
	(267.20, 97.16) --
	(273.38, 95.89) --
	(279.57, 97.03) --
	(285.75, 98.12) --
	(291.93, 99.14) --
	(298.11, 99.59) --
	(304.29,101.00) --
	(310.47,102.50) --
	(316.65,104.29) --
	(322.83,105.29) --
	(329.01,105.81) --
	(335.19,106.66) --
	(341.37,109.14) --
	(347.55,111.98) --
	(353.73,114.58) --
	(359.91,114.85) --
	(366.09,118.06) --
	(372.27,119.90) --
	(378.45,123.28) --
	(384.63,128.74);

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] ( 63.25, 63.80) --
	( 69.44, 63.80) --
	( 75.62, 63.80) --
	( 81.80, 63.80) --
	( 87.98, 63.80) --
	( 94.16, 63.80) --
	(100.34, 63.80) --
	(106.52, 63.80) --
	(112.70, 63.80) --
	(118.88, 63.80) --
	(125.06, 63.80) --
	(131.24, 63.80) --
	(137.42, 63.80) --
	(143.60, 63.80) --
	(149.78, 63.80) --
	(155.96, 63.80) --
	(162.14, 63.80) --
	(168.32, 63.80) --
	(174.50, 63.80) --
	(180.68, 63.80) --
	(186.86, 63.80) --
	(193.04, 63.80) --
	(199.22, 63.80) --
	(205.40, 63.80) --
	(211.58, 63.80) --
	(217.76, 63.80) --
	(223.94, 63.80) --
	(230.12, 63.80) --
	(236.30, 63.80) --
	(242.48, 63.80) --
	(248.66, 63.80) --
	(254.84, 63.80) --
	(261.02, 63.80) --
	(267.20, 63.80) --
	(273.38, 63.80) --
	(279.57, 63.80) --
	(285.75, 63.80) --
	(291.93, 63.80) --
	(298.11, 63.80) --
	(304.29, 63.80) --
	(310.47, 63.80) --
	(316.65, 63.80) --
	(322.83, 63.80) --
	(329.01, 63.80) --
	(335.19, 63.80) --
	(341.37, 63.80) --
	(347.55, 63.80) --
	(353.73, 63.80) --
	(359.91, 63.80) --
	(366.09, 63.80) --
	(372.27, 63.80) --
	(378.45, 63.80) --
	(384.63, 63.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (223.94, 15.60) {Observations used for training sample ($R$)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 12.00, 96.27) {Benchmark};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25, 61.20) -- (372.27, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25, 61.20) -- ( 63.25, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.06, 61.20) -- (125.06, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.86, 61.20) -- (186.86, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.66, 61.20) -- (248.66, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.47, 61.20) -- (310.47, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.27, 61.20) -- (372.27, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 63.25, 39.60) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (125.06, 39.60) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (186.86, 39.60) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (248.66, 39.60) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (310.47, 39.60) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (372.27, 39.60) {70};
\end{scope}
\begin{scope}
\path[clip] ( 50.40, 61.20) rectangle (397.48,131.34);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25, 97.34) --
	( 69.44, 97.88) --
	( 75.62, 98.01) --
	( 81.80, 97.62) --
	( 87.98, 97.87) --
	( 94.16, 98.41) --
	(100.34, 98.76) --
	(106.52, 97.50) --
	(112.70, 97.95) --
	(118.88, 98.43) --
	(125.06, 97.50) --
	(131.24, 97.26) --
	(137.42, 97.93) --
	(143.60, 98.08) --
	(149.78, 98.74) --
	(155.96, 97.99) --
	(162.14, 98.74) --
	(168.32, 99.59) --
	(174.50,100.40) --
	(180.68, 99.28) --
	(186.86,100.08) --
	(193.04,100.89) --
	(199.22, 99.90) --
	(205.40,100.40) --
	(211.58,101.38) --
	(217.76,102.37) --
	(223.94,100.30) --
	(230.12, 94.81) --
	(236.30, 94.33) --
	(242.48, 94.91) --
	(248.66, 94.58) --
	(254.84, 95.42) --
	(261.02, 96.47) --
	(267.20, 97.16) --
	(273.38, 95.89) --
	(279.57, 97.03) --
	(285.75, 98.12) --
	(291.93, 99.14) --
	(298.11, 99.59) --
	(304.29,101.00) --
	(310.47,102.50) --
	(316.65,104.29) --
	(322.83,105.29) --
	(329.01,105.81) --
	(335.19,106.66) --
	(341.37,109.14) --
	(347.55,111.98) --
	(353.73,114.58) --
	(359.91,114.85) --
	(366.09,118.06) --
	(372.27,119.90) --
	(378.45,123.28) --
	(384.63,128.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 63.80) -- ( 50.40,122.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 63.80) -- ( 44.40, 63.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 75.52) -- ( 44.40, 75.52);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 87.23) -- ( 44.40, 87.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 98.95) -- ( 44.40, 98.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40,110.67) -- ( 44.40,110.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40,122.39) -- ( 44.40,122.39);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40, 61.04) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40,119.63) {0.05};
\end{scope}
\end{tikzpicture}