### 2017-04-21 floats/empirics-oos-mse-1b.tex

% Created by tikzDevice version 0.6.2-92-0ad2792 on 2014-03-27 04:41:27
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (397.48,144.54);
\begin{scope}
\path[clip] ( 50.40, 61.20) rectangle (397.48,131.34);
\definecolor[named]{drawColor}{rgb}{1.00,1.00,1.00}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] (248.66, 64.79) --
	(254.84, 71.45) --
	(261.02, 68.74) --
	(267.20, 66.28) --
	(273.38, 63.80) --
	(279.57,105.64) --
	(285.75,104.73) --
	(291.93,113.66) --
	(298.11,116.70) --
	(304.29,115.54) --
	(310.47,113.47) --
	(316.65,120.29) --
	(322.83,121.13) --
	(329.01,119.16) --
	(335.19,121.19) --
	(341.37,120.79) --
	(347.55,121.75) --
	(353.73,120.97) --
	(359.91,125.03) --
	(366.09,124.93) --
	(372.27,120.92) --
	(378.45,115.37) --
	(384.63,110.14);

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] (248.66, 67.69) --
	(254.84, 74.06) --
	(261.02, 71.45) --
	(267.20, 69.09) --
	(273.38, 66.79) --
	(279.57,107.26) --
	(285.75,106.40) --
	(291.93,115.08) --
	(298.11,118.01) --
	(304.29,116.97) --
	(310.47,115.08) --
	(316.65,121.60) --
	(322.83,122.36) --
	(329.01,120.66) --
	(335.19,122.70) --
	(341.37,122.44) --
	(347.55,123.48) --
	(353.73,122.73) --
	(359.91,126.40) --
	(366.09,126.32) --
	(372.27,122.61) --
	(378.45,117.59) --
	(384.63,112.83);

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round] ( 63.25,128.74) --
	( 69.44,128.74) --
	( 75.62,128.74) --
	( 81.80,128.74) --
	( 87.98,128.74) --
	( 94.16,128.74) --
	(100.34,128.74) --
	(106.52,128.74) --
	(112.70,128.74) --
	(118.88,128.74) --
	(125.06,128.74) --
	(131.24,128.74) --
	(137.42,128.74) --
	(143.60,128.74) --
	(149.78,128.74) --
	(155.96,128.74) --
	(162.14,128.74) --
	(168.32,128.74) --
	(174.50,128.74) --
	(180.68,128.74) --
	(186.86,128.74) --
	(193.04,128.74) --
	(199.22,128.74) --
	(205.40,128.74) --
	(211.58,128.74) --
	(217.76,128.74) --
	(223.94,128.74) --
	(230.12,128.74) --
	(236.30,128.74) --
	(242.48,128.74) --
	(248.66,128.74) --
	(254.84,128.74) --
	(261.02,128.74) --
	(267.20,128.74) --
	(273.38,128.74) --
	(279.57,128.74) --
	(285.75,128.74) --
	(291.93,128.74) --
	(298.11,128.74) --
	(304.29,128.74) --
	(310.47,128.74) --
	(316.65,128.74) --
	(322.83,128.74) --
	(329.01,128.74) --
	(335.19,128.74) --
	(341.37,128.74) --
	(347.55,128.74) --
	(353.73,128.74) --
	(359.91,128.74) --
	(366.09,128.74) --
	(372.27,128.74) --
	(378.45,128.74) --
	(384.63,128.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (223.94, 15.60) {Observations used for training sample ($R$)};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 12.00, 96.27) {MSE Diff.};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25, 61.20) -- (372.27, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25, 61.20) -- ( 63.25, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (125.06, 61.20) -- (125.06, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.86, 61.20) -- (186.86, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.66, 61.20) -- (248.66, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.47, 61.20) -- (310.47, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (372.27, 61.20) -- (372.27, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 63.25, 39.60) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (125.06, 39.60) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (186.86, 39.60) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (248.66, 39.60) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (310.47, 39.60) {60};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (372.27, 39.60) {70};
\end{scope}
\begin{scope}
\path[clip] ( 50.40, 61.20) rectangle (397.48,131.34);
\definecolor[named]{fillColor}{rgb}{0.75,0.75,0.75}

\path[fill=fillColor] (248.66,128.74) --
	(248.66, 64.79) --
	(254.84, 71.45) --
	(261.02, 68.74) --
	(267.20, 66.28) --
	(273.38, 63.80) --
	(279.57,105.64) --
	(285.75,104.73) --
	(291.93,113.66) --
	(298.11,116.70) --
	(304.29,115.54) --
	(310.47,113.47) --
	(316.65,120.29) --
	(322.83,121.13) --
	(329.01,119.16) --
	(335.19,121.19) --
	(341.37,120.79) --
	(347.55,121.75) --
	(353.73,120.97) --
	(359.91,125.03) --
	(366.09,124.93) --
	(372.27,120.92) --
	(378.45,115.37) --
	(384.63,110.14) --
	(384.63,128.74) --
	cycle;
\definecolor[named]{drawColor}{rgb}{0.75,0.75,0.75}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.25,128.74) --
	(384.63,128.74);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (248.66, 67.69) --
	(254.84, 74.06) --
	(261.02, 71.45) --
	(267.20, 69.09) --
	(273.38, 66.79) --
	(279.57,107.26) --
	(285.75,106.40) --
	(291.93,115.08) --
	(298.11,118.01) --
	(304.29,116.97) --
	(310.47,115.08) --
	(316.65,121.60) --
	(322.83,122.36) --
	(329.01,120.66) --
	(335.19,122.70) --
	(341.37,122.44) --
	(347.55,123.48) --
	(353.73,122.73) --
	(359.91,126.40) --
	(366.09,126.32) --
	(372.27,122.61) --
	(378.45,117.59) --
	(384.63,112.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,144.54);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 77.95) -- ( 50.40,128.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 77.95) -- ( 44.40, 77.95);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40, 90.65) -- ( 44.40, 90.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40,103.34) -- ( 44.40,103.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40,116.04) -- ( 44.40,116.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.40,128.74) -- ( 44.40,128.74);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40, 75.19) {-.10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40,100.59) {-0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 38.40,125.99) {0};
\end{scope}
\end{tikzpicture}