### 2017-04-21 floats/mc-dmwpower-rec.tex

% Created by tikzDevice version 0.6.2-92-0ad2792 on 2014-03-27 14:58:15
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,325.21);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);

\path[] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (220.94, 12.05) {$P/T$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 14.29,162.76) {Rejection Probability};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 54.31) -- ( 48.75, 54.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 66.59) -- ( 48.75, 66.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 78.87) -- ( 48.75, 78.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 91.15) -- ( 48.75, 91.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,103.43) -- ( 48.75,103.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,115.71) -- ( 48.75,115.71);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 51.00) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 63.28) {0.05};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 75.56) {0.10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 87.85) {0.15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,100.13) {0.20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,112.41) {0.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.20, 50.02) -- ( 57.20, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.76, 50.02) -- ( 80.76, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.31, 50.02) -- (104.31, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.87, 50.02) -- (127.87, 44.32);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 57.20, 32.02) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 80.76, 32.02) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (104.31, 32.02) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (127.87, 32.02) {0.6};
\end{scope}
\begin{scope}
\path[clip] ( 54.44, 50.02) rectangle (137.69,119.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.20, 78.87) --
	(135.72, 78.87);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.56, 78.85) --
	( 59.79, 79.14) --
	( 60.03, 78.68) --
	( 60.26, 79.14) --
	( 60.50, 79.72) --
	( 60.73, 79.32) --
	( 60.97, 78.15) --
	( 61.20, 78.21) --
	( 61.44, 78.39) --
	( 61.68, 79.03) --
	( 61.91, 77.98) --
	( 62.15, 78.04) --
	( 62.38, 78.04) --
	( 62.62, 77.75) --
	( 62.85, 77.98) --
	( 63.09, 77.92) --
	( 63.32, 77.22) --
	( 63.56, 76.99) --
	( 63.80, 76.87) --
	( 64.03, 76.47) --
	( 64.27, 76.12) --
	( 64.50, 75.77) --
	( 64.74, 75.77) --
	( 64.97, 75.07) --
	( 65.21, 74.72) --
	( 65.44, 74.20) --
	( 65.68, 74.14) --
	( 65.92, 73.91) --
	( 66.15, 74.02) --
	( 66.39, 74.14) --
	( 66.62, 74.49) --
	( 66.86, 73.73) --
	( 67.09, 73.68) --
	( 67.33, 73.56) --
	( 67.56, 73.56) --
	( 67.80, 74.31) --
	( 68.04, 74.55) --
	( 68.27, 73.27) --
	( 68.51, 73.09) --
	( 68.74, 72.74) --
	( 68.98, 72.86) --
	( 69.21, 72.40) --
	( 69.45, 72.86) --
	( 69.69, 72.05) --
	( 69.92, 71.58) --
	( 70.16, 72.05) --
	( 70.39, 72.22) --
	( 70.63, 71.87) --
	( 70.86, 71.93) --
	( 71.10, 71.81) --
	( 71.33, 71.29) --
	( 71.57, 70.94) --
	( 71.81, 70.65) --
	( 72.04, 70.65) --
	( 72.28, 71.23) --
	( 72.51, 71.00) --
	( 72.75, 70.71) --
	( 72.98, 70.82) --
	( 73.22, 70.88) --
	( 73.45, 70.24) --
	( 73.69, 70.71) --
	( 73.93, 70.48) --
	( 74.16, 69.78) --
	( 74.40, 70.01) --
	( 74.63, 70.42) --
	( 74.87, 70.36) --
	( 75.10, 70.07) --
	( 75.34, 69.49) --
	( 75.57, 69.20) --
	( 75.81, 69.55) --
	( 76.05, 69.37) --
	( 76.28, 69.66) --
	( 76.52, 69.08) --
	( 76.75, 69.08) --
	( 76.99, 68.96) --
	( 77.22, 69.20) --
	( 77.46, 69.31) --
	( 77.69, 68.73) --
	( 77.93, 68.32) --
	( 78.17, 68.27) --
	( 78.40, 67.80) --
	( 78.64, 68.27) --
	( 78.87, 68.44) --
	( 79.11, 68.21) --
	( 79.34, 68.03) --
	( 79.58, 68.09) --
	( 79.81, 68.15) --
	( 80.05, 67.97) --
	( 80.29, 67.86) --
	( 80.52, 67.92) --
	( 80.76, 68.09) --
	( 80.99, 67.80) --
	( 81.23, 67.28) --
	( 81.46, 67.33) --
	( 81.70, 67.04) --
	( 81.93, 67.63) --
	( 82.17, 67.39) --
	( 82.41, 67.68) --
	( 82.64, 67.39) --
	( 82.88, 67.63) --
	( 83.11, 67.63) --
	( 83.35, 67.10) --
	( 83.58, 66.87) --
	( 83.82, 66.75) --
	( 84.05, 66.75) --
	( 84.29, 66.64) --
	( 84.53, 66.40) --
	( 84.76, 66.35) --
	( 85.00, 66.29) --
	( 85.23, 66.81) --
	( 85.47, 66.58) --
	( 85.70, 66.29) --
	( 85.94, 66.29) --
	( 86.17, 66.17) --
	( 86.41, 66.17) --
	( 86.65, 66.40) --
	( 86.88, 65.65) --
	( 87.12, 65.65) --
	( 87.35, 65.30) --
	( 87.59, 65.47) --
	( 87.82, 65.65) --
	( 88.06, 65.76) --
	( 88.29, 65.65) --
	( 88.53, 65.88) --
	( 88.77, 65.36) --
	( 89.00, 65.36) --
	( 89.24, 65.30) --
	( 89.47, 65.30) --
	( 89.71, 65.18) --
	( 89.94, 65.12) --
	( 90.18, 64.83) --
	( 90.41, 64.60) --
	( 90.65, 64.37) --
	( 90.89, 64.54) --
	( 91.12, 64.37) --
	( 91.36, 64.43) --
	( 91.59, 64.48) --
	( 91.83, 64.14) --
	( 92.06, 64.08) --
	( 92.30, 64.08) --
	( 92.53, 63.50) --
	( 92.77, 63.55) --
	( 93.01, 63.50) --
	( 93.24, 63.55) --
	( 93.48, 63.67) --
	( 93.71, 63.50) --
	( 93.95, 63.38) --
	( 94.18, 63.55) --
	( 94.42, 63.44) --
	( 94.66, 63.44) --
	( 94.89, 63.50) --
	( 95.13, 63.50) --
	( 95.36, 63.55) --
	( 95.60, 63.50) --
	( 95.83, 63.38) --
	( 96.07, 63.26) --
	( 96.30, 63.26) --
	( 96.54, 63.15) --
	( 96.78, 62.80) --
	( 97.01, 63.09) --
	( 97.25, 62.86) --
	( 97.48, 62.57) --
	( 97.72, 62.68) --
	( 97.95, 62.39) --
	( 98.19, 62.39) --
	( 98.42, 62.45) --
	( 98.66, 62.22) --
	( 98.90, 62.10) --
	( 99.13, 62.10) --
	( 99.37, 62.33) --
	( 99.60, 62.39) --
	( 99.84, 62.33) --
	(100.07, 61.93) --
	(100.31, 61.81) --
	(100.54, 61.69) --
	(100.78, 61.52) --
	(101.02, 61.29) --
	(101.25, 61.34) --
	(101.49, 61.58) --
	(101.72, 61.17) --
	(101.96, 60.88) --
	(102.19, 61.05) --
	(102.43, 61.23) --
	(102.66, 61.05) --
	(102.90, 60.82) --
	(103.14, 61.11) --
	(103.37, 61.05) --
	(103.61, 61.29) --
	(103.84, 61.05) --
	(104.08, 61.05) --
	(104.31, 60.94) --
	(104.55, 60.70) --
	(104.78, 60.82) --
	(105.02, 60.88) --
	(105.26, 60.99) --
	(105.49, 60.70) --
	(105.73, 60.70) --
	(105.96, 60.76) --
	(106.20, 60.53) --
	(106.43, 60.47) --
	(106.67, 60.47) --
	(106.90, 60.41) --
	(107.14, 60.30) --
	(107.38, 60.41) --
	(107.61, 60.30) --
	(107.85, 60.12) --
	(108.08, 60.24) --
	(108.32, 60.18) --
	(108.55, 60.12) --
	(108.79, 60.12) --
	(109.02, 60.24) --
	(109.26, 60.18) --
	(109.50, 60.06) --
	(109.73, 59.83) --
	(109.97, 59.89) --
	(110.20, 59.77) --
	(110.44, 59.54) --
	(110.67, 59.37) --
	(110.91, 59.25) --
	(111.14, 59.25) --
	(111.38, 59.13) --
	(111.62, 59.13) --
	(111.85, 59.02) --
	(112.09, 58.96) --
	(112.32, 58.84) --
	(112.56, 58.96) --
	(112.79, 58.96) --
	(113.03, 58.96) --
	(113.26, 58.96) --
	(113.50, 59.02) --
	(113.74, 58.96) --
	(113.97, 58.90) --
	(114.21, 58.84) --
	(114.44, 58.61) --
	(114.68, 58.67) --
	(114.91, 58.49) --
	(115.15, 58.44) --
	(115.38, 58.44) --
	(115.62, 58.38) --
	(115.86, 58.38) --
	(116.09, 58.38) --
	(116.33, 58.32) --
	(116.56, 58.67) --
	(116.80, 58.61) --
	(117.03, 58.26) --
	(117.27, 58.49) --
	(117.51, 58.44) --
	(117.74, 58.38) --
	(117.98, 58.38) --
	(118.21, 58.32) --
	(118.45, 58.20) --
	(118.68, 58.09) --
	(118.92, 58.09) --
	(119.15, 58.14) --
	(119.39, 58.20) --
	(119.63, 58.26) --
	(119.86, 58.26) --
	(120.10, 58.14) --
	(120.33, 57.97) --
	(120.57, 58.03) --
	(120.80, 58.09) --
	(121.04, 57.85) --
	(121.27, 57.97) --
	(121.51, 57.85) --
	(121.75, 57.80) --
	(121.98, 57.80) --
	(122.22, 57.56) --
	(122.45, 57.56) --
	(122.69, 57.45) --
	(122.92, 57.50) --
	(123.16, 57.33) --
	(123.39, 57.39) --
	(123.63, 57.33) --
	(123.87, 57.21) --
	(124.10, 57.27) --
	(124.34, 57.21) --
	(124.57, 56.98) --
	(124.81, 56.92) --
	(125.04, 56.98) --
	(125.28, 56.98) --
	(125.51, 56.81) --
	(125.75, 56.75) --
	(125.99, 56.86) --
	(126.22, 56.75) --
	(126.46, 56.81) --
	(126.69, 56.75) --
	(126.93, 56.75) --
	(127.16, 56.81) --
	(127.40, 56.75) --
	(127.63, 56.63) --
	(127.87, 56.52) --
	(128.11, 56.46) --
	(128.34, 56.46) --
	(128.58, 56.40) --
	(128.81, 56.34) --
	(129.05, 56.40) --
	(129.28, 56.17) --
	(129.52, 56.23) --
	(129.75, 56.11) --
	(129.99, 56.28) --
	(130.23, 56.17) --
	(130.46, 56.11) --
	(130.70, 56.17) --
	(130.93, 56.05) --
	(131.17, 55.93) --
	(131.40, 56.05) --
	(131.64, 55.88) --
	(131.87, 55.93) --
	(132.11, 55.99) --
	(132.35, 55.82) --
	(132.58, 55.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 50.02) rectangle (137.69,119.88);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,119.88) rectangle (137.69,127.83);

\path[] ( 54.44,119.88) rectangle (137.69,127.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 69.67,121.58) {T=500, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,119.88) rectangle (137.69,127.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.45, 50.02) -- (140.45, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.01, 50.02) -- (164.01, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.56, 50.02) -- (187.56, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.12, 50.02) -- (211.12, 44.32);
\end{scope}
\begin{scope}
\path[clip] (137.69, 50.02) rectangle (220.94,119.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.45, 78.87) --
	(218.97, 78.87);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (142.80, 75.31) --
	(143.04, 74.32) --
	(143.28, 75.80) --
	(143.51, 75.80) --
	(143.75, 75.06) --
	(143.98, 75.18) --
	(144.22, 74.69) --
	(144.45, 74.32) --
	(144.69, 74.20) --
	(144.92, 73.10) --
	(145.16, 72.97) --
	(145.40, 71.75) --
	(145.63, 71.62) --
	(145.87, 71.38) --
	(146.10, 72.97) --
	(146.34, 74.45) --
	(146.57, 73.71) --
	(146.81, 73.34) --
	(147.05, 71.13) --
	(147.28, 69.90) --
	(147.52, 70.64) --
	(147.75, 71.62) --
	(147.99, 71.13) --
	(148.22, 71.13) --
	(148.46, 71.62) --
	(148.69, 70.64) --
	(148.93, 70.15) --
	(149.17, 70.76) --
	(149.40, 70.39) --
	(149.64, 71.13) --
	(149.87, 71.75) --
	(150.11, 69.66) --
	(150.34, 69.78) --
	(150.58, 69.66) --
	(150.81, 68.80) --
	(151.05, 69.29) --
	(151.29, 69.29) --
	(151.52, 68.92) --
	(151.76, 68.92) --
	(151.99, 68.92) --
	(152.23, 67.57) --
	(152.46, 68.18) --
	(152.70, 67.20) --
	(152.93, 67.69) --
	(153.17, 67.69) --
	(153.41, 68.18) --
	(153.64, 68.68) --
	(153.88, 68.55) --
	(154.11, 68.43) --
	(154.35, 68.43) --
	(154.58, 67.69) --
	(154.82, 67.32) --
	(155.05, 67.94) --
	(155.29, 68.06) --
	(155.53, 67.45) --
	(155.76, 67.20) --
	(156.00, 66.83) --
	(156.23, 66.34) --
	(156.47, 66.34) --
	(156.70, 66.71) --
	(156.94, 66.34) --
	(157.17, 66.59) --
	(157.41, 65.85) --
	(157.65, 65.60) --
	(157.88, 66.34) --
	(158.12, 65.48) --
	(158.35, 65.36) --
	(158.59, 65.85) --
	(158.82, 65.36) --
	(159.06, 66.59) --
	(159.29, 65.97) --
	(159.53, 65.60) --
	(159.77, 66.22) --
	(160.00, 65.85) --
	(160.24, 65.85) --
	(160.47, 64.99) --
	(160.71, 64.99) --
	(160.94, 65.60) --
	(161.18, 65.85) --
	(161.41, 65.11) --
	(161.65, 64.25) --
	(161.89, 63.89) --
	(162.12, 64.25) --
	(162.36, 64.01) --
	(162.59, 64.13) --
	(162.83, 64.75) --
	(163.06, 64.13) --
	(163.30, 64.13) --
	(163.53, 64.13) --
	(163.77, 64.38) --
	(164.01, 64.25) --
	(164.24, 64.01) --
	(164.48, 63.64) --
	(164.71, 63.64) --
	(164.95, 63.52) --
	(165.18, 63.76) --
	(165.42, 63.76) --
	(165.65, 63.27) --
	(165.89, 63.15) --
	(166.13, 63.76) --
	(166.36, 63.89) --
	(166.60, 63.03) --
	(166.83, 62.66) --
	(167.07, 62.29) --
	(167.30, 62.66) --
	(167.54, 61.92) --
	(167.77, 61.92) --
	(168.01, 61.67) --
	(168.25, 61.80) --
	(168.48, 61.92) --
	(168.72, 62.41) --
	(168.95, 62.41) --
	(169.19, 62.53) --
	(169.42, 62.41) --
	(169.66, 62.17) --
	(169.90, 61.80) --
	(170.13, 61.67) --
	(170.37, 61.80) --
	(170.60, 61.67) --
	(170.84, 61.80) --
	(171.07, 61.80) --
	(171.31, 60.94) --
	(171.54, 60.45) --
	(171.78, 60.32) --
	(172.02, 60.45) --
	(172.25, 60.20) --
	(172.49, 60.57) --
	(172.72, 61.18) --
	(172.96, 60.57) --
	(173.19, 60.08) --
	(173.43, 60.20) --
	(173.66, 60.08) --
	(173.90, 60.08) --
	(174.14, 60.08) --
	(174.37, 60.08) --
	(174.61, 59.34) --
	(174.84, 59.22) --
	(175.08, 59.22) --
	(175.31, 58.97) --
	(175.55, 58.73) --
	(175.78, 58.85) --
	(176.02, 58.85) --
	(176.26, 59.22) --
	(176.49, 59.22) --
	(176.73, 59.10) --
	(176.96, 59.10) --
	(177.20, 59.34) --
	(177.43, 58.85) --
	(177.67, 58.85) --
	(177.90, 58.85) --
	(178.14, 58.73) --
	(178.38, 58.36) --
	(178.61, 58.48) --
	(178.85, 57.99) --
	(179.08, 58.11) --
	(179.32, 57.99) --
	(179.55, 57.87) --
	(179.79, 57.74) --
	(180.02, 57.38) --
	(180.26, 57.25) --
	(180.50, 57.13) --
	(180.73, 57.50) --
	(180.97, 57.38) --
	(181.20, 57.13) --
	(181.44, 57.25) --
	(181.67, 57.62) --
	(181.91, 57.25) --
	(182.14, 57.25) --
	(182.38, 57.13) --
	(182.62, 56.76) --
	(182.85, 56.88) --
	(183.09, 56.52) --
	(183.32, 56.27) --
	(183.56, 56.27) --
	(183.79, 56.39) --
	(184.03, 56.15) --
	(184.26, 56.03) --
	(184.50, 56.39) --
	(184.74, 56.03) --
	(184.97, 56.27) --
	(185.21, 56.15) --
	(185.44, 56.27) --
	(185.68, 56.15) --
	(185.91, 56.39) --
	(186.15, 56.27) --
	(186.38, 56.52) --
	(186.62, 56.27) --
	(186.86, 56.15) --
	(187.09, 56.15) --
	(187.33, 55.90) --
	(187.56, 55.78) --
	(187.80, 55.53) --
	(188.03, 55.78) --
	(188.27, 55.66) --
	(188.50, 55.41) --
	(188.74, 55.29) --
	(188.98, 55.41) --
	(189.21, 55.66) --
	(189.45, 55.53) --
	(189.68, 55.29) --
	(189.92, 54.92) --
	(190.15, 55.04) --
	(190.39, 55.29) --
	(190.62, 55.41) --
	(190.86, 55.53) --
	(191.10, 55.29) --
	(191.33, 55.29) --
	(191.57, 55.04) --
	(191.80, 55.17) --
	(192.04, 55.29) --
	(192.27, 55.29) --
	(192.51, 55.04) --
	(192.74, 55.04) --
	(192.98, 55.04) --
	(193.22, 55.04) --
	(193.45, 55.17) --
	(193.69, 55.17) --
	(193.92, 55.29) --
	(194.16, 55.04) --
	(194.39, 54.92) --
	(194.63, 54.92) --
	(194.87, 54.92) --
	(195.10, 54.92) --
	(195.34, 54.67) --
	(195.57, 54.55) --
	(195.81, 54.55) --
	(196.04, 54.55) --
	(196.28, 54.55) --
	(196.51, 54.55) --
	(196.75, 54.55) --
	(196.99, 54.55) --
	(197.22, 54.55) --
	(197.46, 54.67) --
	(197.69, 54.80) --
	(197.93, 54.80) --
	(198.16, 54.92) --
	(198.40, 54.67) --
	(198.63, 54.80) --
	(198.87, 54.80) --
	(199.11, 54.67) --
	(199.34, 54.43) --
	(199.58, 54.55) --
	(199.81, 54.55) --
	(200.05, 54.55) --
	(200.28, 54.67) --
	(200.52, 54.55) --
	(200.75, 54.55) --
	(200.99, 54.55) --
	(201.23, 54.43) --
	(201.46, 54.55) --
	(201.70, 54.43) --
	(201.93, 54.55) --
	(202.17, 54.55) --
	(202.40, 54.43) --
	(202.64, 54.31) --
	(202.87, 54.31) --
	(203.11, 54.31) --
	(203.35, 54.31) --
	(203.58, 54.31) --
	(203.82, 54.31) --
	(204.05, 54.43) --
	(204.29, 54.55) --
	(204.52, 54.55) --
	(204.76, 54.43) --
	(204.99, 54.43) --
	(205.23, 54.43) --
	(205.47, 54.43) --
	(205.70, 54.43) --
	(205.94, 54.55) --
	(206.17, 54.55) --
	(206.41, 54.55) --
	(206.64, 54.31) --
	(206.88, 54.43) --
	(207.11, 54.55) --
	(207.35, 54.55) --
	(207.59, 54.55) --
	(207.82, 54.55) --
	(208.06, 54.55) --
	(208.29, 54.55) --
	(208.53, 54.43) --
	(208.76, 54.43) --
	(209.00, 54.43) --
	(209.23, 54.31) --
	(209.47, 54.43) --
	(209.71, 54.31) --
	(209.94, 54.31) --
	(210.18, 54.31) --
	(210.41, 54.43) --
	(210.65, 54.43) --
	(210.88, 54.31) --
	(211.12, 54.43) --
	(211.35, 54.31) --
	(211.59, 54.31) --
	(211.83, 54.31) --
	(212.06, 54.31) --
	(212.30, 54.31) --
	(212.53, 54.31) --
	(212.77, 54.31) --
	(213.00, 54.31) --
	(213.24, 54.31) --
	(213.47, 54.31) --
	(213.71, 54.31) --
	(213.95, 54.31) --
	(214.18, 54.31) --
	(214.42, 54.31) --
	(214.65, 54.31) --
	(214.89, 54.31) --
	(215.12, 54.31) --
	(215.36, 54.31) --
	(215.59, 54.31) --
	(215.83, 54.31);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.69, 50.02) rectangle (220.94,119.88);
\end{scope}
\begin{scope}
\path[clip] (137.69,119.88) rectangle (220.94,127.83);

\path[] (137.69,119.88) rectangle (220.94,127.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (147.24,121.58) {T=500, K=T/10, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.69,119.88) rectangle (220.94,127.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.70, 50.02) -- (223.70, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.26, 50.02) -- (247.26, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (270.81, 50.02) -- (270.81, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (294.37, 50.02) -- (294.37, 44.32);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (223.70, 32.02) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (247.26, 32.02) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (270.81, 32.02) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (294.37, 32.02) {0.6};
\end{scope}
\begin{scope}
\path[clip] (220.94, 50.02) rectangle (304.19,119.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.70, 78.87) --
	(302.22, 78.87);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.05, 81.08) --
	(226.29, 80.59) --
	(226.53, 83.54) --
	(226.76, 81.20) --
	(227.00, 81.94) --
	(227.23, 81.08) --
	(227.47, 81.45) --
	(227.70, 81.33) --
	(227.94, 82.55) --
	(228.17, 83.17) --
	(228.41, 82.43) --
	(228.65, 81.69) --
	(228.88, 83.29) --
	(229.12, 81.82) --
	(229.35, 83.05) --
	(229.59, 81.94) --
	(229.82, 81.33) --
	(230.06, 80.71) --
	(230.29, 80.71) --
	(230.53, 81.08) --
	(230.77, 82.19) --
	(231.00, 82.19) --
	(231.24, 82.06) --
	(231.47, 82.43) --
	(231.71, 82.06) --
	(231.94, 82.55) --
	(232.18, 81.94) --
	(232.41, 81.94) --
	(232.65, 80.71) --
	(232.89, 80.47) --
	(233.12, 80.59) --
	(233.36, 81.94) --
	(233.59, 81.45) --
	(233.83, 79.73) --
	(234.06, 79.73) --
	(234.30, 78.62) --
	(234.53, 80.47) --
	(234.77, 80.59) --
	(235.01, 80.96) --
	(235.24, 80.34) --
	(235.48, 80.47) --
	(235.71, 79.85) --
	(235.95, 79.97) --
	(236.18, 80.22) --
	(236.42, 79.11) --
	(236.65, 79.73) --
	(236.89, 78.87) --
	(237.13, 79.85) --
	(237.36, 79.73) --
	(237.60, 78.87) --
	(237.83, 78.87) --
	(238.07, 79.73) --
	(238.30, 79.24) --
	(238.54, 78.62) --
	(238.77, 78.87) --
	(239.01, 80.10) --
	(239.25, 79.97) --
	(239.48, 80.59) --
	(239.72, 80.34) --
	(239.95, 81.69) --
	(240.19, 81.45) --
	(240.42, 80.96) --
	(240.66, 80.96) --
	(240.89, 80.71) --
	(241.13, 80.59) --
	(241.37, 81.08) --
	(241.60, 80.59) --
	(241.84, 80.59) --
	(242.07, 80.22) --
	(242.31, 80.34) --
	(242.54, 80.47) --
	(242.78, 80.34) --
	(243.01, 80.10) --
	(243.25, 80.47) --
	(243.49, 79.85) --
	(243.72, 79.97) --
	(243.96, 80.22) --
	(244.19, 80.59) --
	(244.43, 81.08) --
	(244.66, 81.33) --
	(244.90, 81.57) --
	(245.13, 80.47) --
	(245.37, 79.97) --
	(245.61, 80.10) --
	(245.84, 80.34) --
	(246.08, 80.34) --
	(246.31, 79.97) --
	(246.55, 80.59) --
	(246.78, 80.59) --
	(247.02, 80.22) --
	(247.26, 79.48) --
	(247.49, 78.50) --
	(247.73, 78.01) --
	(247.96, 77.89) --
	(248.20, 77.64) --
	(248.43, 78.01) --
	(248.67, 78.26) --
	(248.90, 78.99) --
	(249.14, 79.24) --
	(249.38, 78.99) --
	(249.61, 78.50) --
	(249.85, 78.13) --
	(250.08, 78.01) --
	(250.32, 78.50) --
	(250.55, 78.13) --
	(250.79, 78.01) --
	(251.02, 78.26) --
	(251.26, 78.62) --
	(251.50, 78.87) --
	(251.73, 78.75) --
	(251.97, 78.38) --
	(252.20, 78.87) --
	(252.44, 78.26) --
	(252.67, 78.62) --
	(252.91, 77.15) --
	(253.14, 77.40) --
	(253.38, 77.64) --
	(253.62, 77.89) --
	(253.85, 77.40) --
	(254.09, 77.40) --
	(254.32, 77.15) --
	(254.56, 77.52) --
	(254.79, 76.78) --
	(255.03, 76.78) --
	(255.26, 76.66) --
	(255.50, 77.52) --
	(255.74, 77.52) --
	(255.97, 77.40) --
	(256.21, 77.03) --
	(256.44, 76.78) --
	(256.68, 76.90) --
	(256.91, 77.40) --
	(257.15, 77.64) --
	(257.38, 77.64) --
	(257.62, 77.89) --
	(257.86, 77.52) --
	(258.09, 76.90) --
	(258.33, 77.40) --
	(258.56, 76.54) --
	(258.80, 76.78) --
	(259.03, 77.15) --
	(259.27, 76.78) --
	(259.50, 76.66) --
	(259.74, 76.54) --
	(259.98, 76.17) --
	(260.21, 76.29) --
	(260.45, 75.43) --
	(260.68, 75.80) --
	(260.92, 76.17) --
	(261.15, 76.04) --
	(261.39, 75.68) --
	(261.62, 75.43) --
	(261.86, 75.68) --
	(262.10, 75.68) --
	(262.33, 75.31) --
	(262.57, 75.43) --
	(262.80, 75.18) --
	(263.04, 75.06) --
	(263.27, 74.94) --
	(263.51, 75.18) --
	(263.74, 74.45) --
	(263.98, 74.69) --
	(264.22, 74.08) --
	(264.45, 73.96) --
	(264.69, 74.45) --
	(264.92, 74.57) --
	(265.16, 74.08) --
	(265.39, 73.71) --
	(265.63, 73.71) --
	(265.86, 73.47) --
	(266.10, 73.83) --
	(266.34, 73.96) --
	(266.57, 73.71) --
	(266.81, 73.10) --
	(267.04, 73.59) --
	(267.28, 73.83) --
	(267.51, 73.96) --
	(267.75, 74.32) --
	(267.98, 74.32) --
	(268.22, 73.96) --
	(268.46, 73.47) --
	(268.69, 73.83) --
	(268.93, 73.71) --
	(269.16, 73.71) --
	(269.40, 73.47) --
	(269.63, 73.59) --
	(269.87, 73.47) --
	(270.11, 72.97) --
	(270.34, 72.73) --
	(270.58, 72.85) --
	(270.81, 72.48) --
	(271.05, 72.73) --
	(271.28, 72.85) --
	(271.52, 72.61) --
	(271.75, 72.36) --
	(271.99, 72.48) --
	(272.23, 72.36) --
	(272.46, 72.11) --
	(272.70, 72.24) --
	(272.93, 71.25) --
	(273.17, 71.25) --
	(273.40, 71.01) --
	(273.64, 71.01) --
	(273.87, 71.25) --
	(274.11, 71.38) --
	(274.35, 71.01) --
	(274.58, 70.39) --
	(274.82, 70.64) --
	(275.05, 71.13) --
	(275.29, 70.76) --
	(275.52, 70.52) --
	(275.76, 70.64) --
	(275.99, 70.27) --
	(276.23, 70.76) --
	(276.47, 70.64) --
	(276.70, 70.64) --
	(276.94, 71.38) --
	(277.17, 71.01) --
	(277.41, 71.01) --
	(277.64, 71.13) --
	(277.88, 71.38) --
	(278.11, 71.01) --
	(278.35, 70.76) --
	(278.59, 70.27) --
	(278.82, 70.64) --
	(279.06, 70.76) --
	(279.29, 71.25) --
	(279.53, 70.39) --
	(279.76, 70.76) --
	(280.00, 71.13) --
	(280.23, 70.39) --
	(280.47, 70.76) --
	(280.71, 70.76) --
	(280.94, 70.39) --
	(281.18, 70.27) --
	(281.41, 70.52) --
	(281.65, 70.39) --
	(281.88, 70.39) --
	(282.12, 70.52) --
	(282.35, 71.01) --
	(282.59, 70.52) --
	(282.83, 70.52) --
	(283.06, 69.90) --
	(283.30, 70.39) --
	(283.53, 69.90) --
	(283.77, 70.27) --
	(284.00, 70.39) --
	(284.24, 70.39) --
	(284.47, 70.27) --
	(284.71, 70.52) --
	(284.95, 70.03) --
	(285.18, 70.64) --
	(285.42, 70.27) --
	(285.65, 70.03) --
	(285.89, 69.90) --
	(286.12, 70.03) --
	(286.36, 70.03) --
	(286.59, 70.52) --
	(286.83, 70.27) --
	(287.07, 70.64) --
	(287.30, 70.64) --
	(287.54, 71.13) --
	(287.77, 70.89) --
	(288.01, 71.01) --
	(288.24, 70.39) --
	(288.48, 70.52) --
	(288.71, 70.89) --
	(288.95, 70.64) --
	(289.19, 70.39) --
	(289.42, 70.27) --
	(289.66, 69.66) --
	(289.89, 69.54) --
	(290.13, 69.41) --
	(290.36, 69.04) --
	(290.60, 69.17) --
	(290.83, 68.92) --
	(291.07, 68.80) --
	(291.31, 68.55) --
	(291.54, 68.68) --
	(291.78, 68.92) --
	(292.01, 69.04) --
	(292.25, 68.55) --
	(292.48, 68.06) --
	(292.72, 67.82) --
	(292.96, 67.57) --
	(293.19, 67.69) --
	(293.43, 67.32) --
	(293.66, 66.96) --
	(293.90, 66.96) --
	(294.13, 66.83) --
	(294.37, 66.83) --
	(294.60, 66.59) --
	(294.84, 66.34) --
	(295.08, 66.96) --
	(295.31, 66.83) --
	(295.55, 66.96) --
	(295.78, 67.08) --
	(296.02, 67.08) --
	(296.25, 66.83) --
	(296.49, 66.71) --
	(296.72, 66.96) --
	(296.96, 66.34) --
	(297.20, 66.46) --
	(297.43, 66.22) --
	(297.67, 66.34) --
	(297.90, 66.46) --
	(298.14, 66.34) --
	(298.37, 66.34) --
	(298.61, 66.10) --
	(298.84, 66.34) --
	(299.08, 66.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.94, 50.02) rectangle (304.19,119.88);
\end{scope}
\begin{scope}
\path[clip] (220.94,119.88) rectangle (304.19,127.83);

\path[] (220.94,119.88) rectangle (304.19,127.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (236.17,121.58) {T=500, K=3, c=2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.94,119.88) rectangle (304.19,127.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (306.95, 50.02) -- (306.95, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.50, 50.02) -- (330.50, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.06, 50.02) -- (354.06, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.62, 50.02) -- (377.62, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44, 54.31) -- (393.13, 54.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44, 66.59) -- (393.13, 66.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44, 78.87) -- (393.13, 78.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44, 91.15) -- (393.13, 91.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,103.43) -- (393.13,103.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,115.71) -- (393.13,115.71);
\end{scope}
\begin{scope}
\path[clip] (304.19, 50.02) rectangle (387.44,119.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (306.95, 78.87) --
	(385.47, 78.87);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (309.30, 89.92) --
	(309.54, 90.29) --
	(309.77, 90.17) --
	(310.01, 90.29) --
	(310.25, 89.31) --
	(310.48, 90.29) --
	(310.72, 91.64) --
	(310.95, 90.54) --
	(311.19, 91.52) --
	(311.42, 91.77) --
	(311.66, 92.62) --
	(311.89, 94.96) --
	(312.13, 96.06) --
	(312.37, 94.71) --
	(312.60, 96.43) --
	(312.84, 97.41) --
	(313.07, 98.52) --
	(313.31, 99.13) --
	(313.54, 98.77) --
	(313.78,100.49) --
	(314.01,100.36) --
	(314.25,100.61) --
	(314.49,101.35) --
	(314.72,102.33) --
	(314.96,101.59) --
	(315.19,101.10) --
	(315.43,101.10) --
	(315.66,100.61) --
	(315.90,102.20) --
	(316.13,102.33) --
	(316.37,102.33) --
	(316.61,102.82) --
	(316.84,104.17) --
	(317.08,104.17) --
	(317.31,103.68) --
	(317.55,104.91) --
	(317.78,103.68) --
	(318.02,104.54) --
	(318.25,105.03) --
	(318.49,105.28) --
	(318.73,105.40) --
	(318.96,105.03) --
	(319.20,104.78) --
	(319.43,103.80) --
	(319.67,104.91) --
	(319.90,106.87) --
	(320.14,107.49) --
	(320.37,107.49) --
	(320.61,107.61) --
	(320.85,107.98) --
	(321.08,107.98) --
	(321.32,108.10) --
	(321.55,108.47) --
	(321.79,107.98) --
	(322.02,108.35) --
	(322.26,108.71) --
	(322.50,107.98) --
	(322.73,109.21) --
	(322.97,108.22) --
	(323.20,108.71) --
	(323.44,109.70) --
	(323.67,109.70) --
	(323.91,110.68) --
	(324.14,109.94) --
	(324.38,110.31) --
	(324.62,109.33) --
	(324.85,108.84) --
	(325.09,110.07) --
	(325.32,109.45) --
	(325.56,109.57) --
	(325.79,110.80) --
	(326.03,110.43) --
	(326.26,110.31) --
	(326.50,111.78) --
	(326.74,110.68) --
	(326.97,111.29) --
	(327.21,111.91) --
	(327.44,111.29) --
	(327.68,111.17) --
	(327.91,110.56) --
	(328.15,110.31) --
	(328.38,110.80) --
	(328.62,110.43) --
	(328.86,111.78) --
	(329.09,111.42) --
	(329.33,111.66) --
	(329.56,113.01) --
	(329.80,112.64) --
	(330.03,112.77) --
	(330.27,112.89) --
	(330.50,113.87) --
	(330.74,114.12) --
	(330.98,112.15) --
	(331.21,111.54) --
	(331.45,113.50) --
	(331.68,113.50) --
	(331.92,114.98) --
	(332.15,115.22) --
	(332.39,115.59) --
	(332.62,114.73) --
	(332.86,115.35) --
	(333.10,113.75) --
	(333.33,114.24) --
	(333.57,113.50) --
	(333.80,114.98) --
	(334.04,115.10) --
	(334.27,113.87) --
	(334.51,114.24) --
	(334.74,114.36) --
	(334.98,114.24) --
	(335.22,113.01) --
	(335.45,113.87) --
	(335.69,113.87) --
	(335.92,114.61) --
	(336.16,115.22) --
	(336.39,114.00) --
	(336.63,114.36) --
	(336.86,113.75) --
	(337.10,114.61) --
	(337.34,113.38) --
	(337.57,112.89) --
	(337.81,112.77) --
	(338.04,113.14) --
	(338.28,112.03) --
	(338.51,111.66) --
	(338.75,112.77) --
	(338.98,112.52) --
	(339.22,111.66) --
	(339.46,111.66) --
	(339.69,111.42) --
	(339.93,111.17) --
	(340.16,112.28) --
	(340.40,111.29) --
	(340.63,109.45) --
	(340.87,110.31) --
	(341.10,110.43) --
	(341.34,111.17) --
	(341.58,110.80) --
	(341.81,110.92) --
	(342.05,111.29) --
	(342.28,110.80) --
	(342.52,111.17) --
	(342.75,110.92) --
	(342.99,111.66) --
	(343.22,111.54) --
	(343.46,111.54) --
	(343.70,110.92) --
	(343.93,111.54) --
	(344.17,110.92) --
	(344.40,111.17) --
	(344.64,111.05) --
	(344.87,110.68) --
	(345.11,109.82) --
	(345.35,110.92) --
	(345.58,110.68) --
	(345.82,111.54) --
	(346.05,110.56) --
	(346.29,110.19) --
	(346.52,110.80) --
	(346.76,110.92) --
	(346.99,111.29) --
	(347.23,111.05) --
	(347.47,110.68) --
	(347.70,110.19) --
	(347.94,110.31) --
	(348.17,110.19) --
	(348.41,110.07) --
	(348.64,110.31) --
	(348.88,109.70) --
	(349.11,108.71) --
	(349.35,109.57) --
	(349.59,108.96) --
	(349.82,109.57) --
	(350.06,109.45) --
	(350.29,109.45) --
	(350.53,108.35) --
	(350.76,107.85) --
	(351.00,107.85) --
	(351.23,108.35) --
	(351.47,108.96) --
	(351.71,108.96) --
	(351.94,107.85) --
	(352.18,108.84) --
	(352.41,108.47) --
	(352.65,108.71) --
	(352.88,107.98) --
	(353.12,107.49) --
	(353.35,106.63) --
	(353.59,106.99) --
	(353.83,105.15) --
	(354.06,105.40) --
	(354.30,104.42) --
	(354.53,103.68) --
	(354.77,103.43) --
	(355.00,103.80) --
	(355.24,104.17) --
	(355.47,103.56) --
	(355.71,103.19) --
	(355.95,103.31) --
	(356.18,103.06) --
	(356.42,102.45) --
	(356.65,101.59) --
	(356.89,101.84) --
	(357.12,102.08) --
	(357.36,101.22) --
	(357.59,102.20) --
	(357.83,101.71) --
	(358.07,101.71) --
	(358.30,100.73) --
	(358.54,100.24) --
	(358.77,101.47) --
	(359.01,101.10) --
	(359.24, 99.50) --
	(359.48, 99.38) --
	(359.71, 99.38) --
	(359.95, 99.75) --
	(360.19,100.85) --
	(360.42, 99.26) --
	(360.66, 99.01) --
	(360.89, 98.52) --
	(361.13, 98.52) --
	(361.36, 98.03) --
	(361.60, 96.80) --
	(361.83, 97.41) --
	(362.07, 95.94) --
	(362.31, 95.70) --
	(362.54, 96.80) --
	(362.78, 96.80) --
	(363.01, 96.06) --
	(363.25, 95.70) --
	(363.48, 95.20) --
	(363.72, 94.84) --
	(363.95, 95.20) --
	(364.19, 95.08) --
	(364.43, 93.85) --
	(364.66, 94.34) --
	(364.90, 94.10) --
	(365.13, 94.59) --
	(365.37, 93.48) --
	(365.60, 93.73) --
	(365.84, 94.22) --
	(366.07, 93.36) --
	(366.31, 93.36) --
	(366.55, 93.12) --
	(366.78, 93.24) --
	(367.02, 93.24) --
	(367.25, 93.61) --
	(367.49, 93.61) --
	(367.72, 92.26) --
	(367.96, 92.01) --
	(368.19, 90.41) --
	(368.43, 91.15) --
	(368.67, 90.91) --
	(368.90, 91.27) --
	(369.14, 91.03) --
	(369.37, 91.89) --
	(369.61, 91.27) --
	(369.84, 91.40) --
	(370.08, 91.03) --
	(370.32, 90.05) --
	(370.55, 89.80) --
	(370.79, 89.55) --
	(371.02, 89.19) --
	(371.26, 89.19) --
	(371.49, 88.45) --
	(371.73, 87.96) --
	(371.96, 87.96) --
	(372.20, 87.47) --
	(372.44, 87.84) --
	(372.67, 87.10) --
	(372.91, 86.98) --
	(373.14, 86.48) --
	(373.38, 85.50) --
	(373.61, 85.62) --
	(373.85, 84.40) --
	(374.08, 85.01) --
	(374.32, 84.64) --
	(374.56, 84.76) --
	(374.79, 84.89) --
	(375.03, 83.66) --
	(375.26, 83.90) --
	(375.50, 83.78) --
	(375.73, 83.41) --
	(375.97, 82.92) --
	(376.20, 81.82) --
	(376.44, 81.69) --
	(376.68, 81.20) --
	(376.91, 81.08) --
	(377.15, 80.71) --
	(377.38, 80.47) --
	(377.62, 79.48) --
	(377.85, 78.50) --
	(378.09, 78.99) --
	(378.32, 79.24) --
	(378.56, 78.87) --
	(378.80, 78.50) --
	(379.03, 78.26) --
	(379.27, 78.01) --
	(379.50, 77.52) --
	(379.74, 76.90) --
	(379.97, 77.15) --
	(380.21, 76.66) --
	(380.44, 76.66) --
	(380.68, 76.29) --
	(380.92, 76.29) --
	(381.15, 75.80) --
	(381.39, 75.31) --
	(381.62, 75.18) --
	(381.86, 74.94) --
	(382.09, 75.18) --
	(382.33, 74.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.19, 50.02) rectangle (387.44,119.88);
\end{scope}
\begin{scope}
\path[clip] (304.19,119.88) rectangle (387.44,127.83);

\path[] (304.19,119.88) rectangle (387.44,127.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (313.74,121.58) {T=500, K=T/10, c=2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.19,119.88) rectangle (387.44,127.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,132.12) -- ( 48.75,132.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,144.40) -- ( 48.75,144.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,156.69) -- ( 48.75,156.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,168.97) -- ( 48.75,168.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,181.25) -- ( 48.75,181.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,193.53) -- ( 48.75,193.53);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,127.83) rectangle (137.69,197.70);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.20,156.69) --
	(135.72,156.69);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.91,156.44) --
	( 62.38,157.24) --
	( 62.85,157.35) --
	( 63.32,157.18) --
	( 63.80,157.47) --
	( 64.27,156.21) --
	( 64.74,156.38) --
	( 65.21,156.27) --
	( 65.68,156.38) --
	( 66.15,156.67) --
	( 66.62,155.87) --
	( 67.09,156.10) --
	( 67.56,155.53) --
	( 68.04,154.50) --
	( 68.51,154.62) --
	( 68.98,154.79) --
	( 69.45,154.39) --
	( 69.92,154.11) --
	( 70.39,154.16) --
	( 70.86,153.99) --
	( 71.33,153.71) --
	( 71.81,153.48) --
	( 72.28,153.42) --
	( 72.75,152.74) --
	( 73.22,152.23) --
	( 73.69,152.45) --
	( 74.16,153.14) --
	( 74.63,151.83) --
	( 75.10,151.66) --
	( 75.57,151.49) --
	( 76.05,151.49) --
	( 76.52,151.37) --
	( 76.99,151.03) --
	( 77.46,150.57) --
	( 77.93,150.63) --
	( 78.40,150.57) --
	( 78.87,150.00) --
	( 79.34,149.66) --
	( 79.81,150.00) --
	( 80.29,150.06) --
	( 80.76,149.38) --
	( 81.23,149.49) --
	( 81.70,148.87) --
	( 82.17,149.04) --
	( 82.64,149.09) --
	( 83.11,148.52) --
	( 83.58,148.24) --
	( 84.05,148.47) --
	( 84.53,148.41) --
	( 85.00,147.56) --
	( 85.47,147.27) --
	( 85.94,147.16) --
	( 86.41,146.64) --
	( 86.88,146.19) --
	( 87.35,146.36) --
	( 87.82,145.62) --
	( 88.29,145.90) --
	( 88.77,145.79) --
	( 89.24,146.19) --
	( 89.71,145.68) --
	( 90.18,144.82) --
	( 90.65,144.71) --
	( 91.12,145.11) --
	( 91.59,144.82) --
	( 92.06,144.59) --
	( 92.53,144.59) --
	( 93.01,143.97) --
	( 93.48,143.63) --
	( 93.95,143.17) --
	( 94.42,143.11) --
	( 94.89,143.28) --
	( 95.36,143.46) --
	( 95.83,143.23) --
	( 96.30,143.46) --
	( 96.78,143.40) --
	( 97.25,142.89) --
	( 97.72,142.71) --
	( 98.19,142.66) --
	( 98.66,142.20) --
	( 99.13,141.97) --
	( 99.60,141.97) --
	(100.07,141.75) --
	(100.54,141.52) --
	(101.02,141.41) --
	(101.49,141.12) --
	(101.96,141.06) --
	(102.43,140.89) --
	(102.90,140.84) --
	(103.37,140.61) --
	(103.84,140.55) --
	(104.31,140.84) --
	(104.78,140.78) --
	(105.26,140.84) --
	(105.73,140.61) --
	(106.20,140.38) --
	(106.67,140.78) --
	(107.14,140.49) --
	(107.61,140.38) --
	(108.08,140.27) --
	(108.55,140.10) --
	(109.02,139.87) --
	(109.50,139.92) --
	(109.97,139.70) --
	(110.44,139.35) --
	(110.91,139.07) --
	(111.38,138.61) --
	(111.85,138.61) --
	(112.32,138.56) --
	(112.79,138.61) --
	(113.26,138.44) --
	(113.74,138.27) --
	(114.21,138.04) --
	(114.68,138.22) --
	(115.15,138.33) --
	(115.62,137.87) --
	(116.09,137.99) --
	(116.56,137.65) --
	(117.03,137.59) --
	(117.51,137.48) --
	(117.98,137.59) --
	(118.45,137.42) --
	(118.92,137.36) --
	(119.39,136.96) --
	(119.86,136.68) --
	(120.33,136.68) --
	(120.80,136.62) --
	(121.27,136.34) --
	(121.75,135.94) --
	(122.22,136.05) --
	(122.69,135.71) --
	(123.16,135.65) --
	(123.63,135.71) --
	(124.10,135.60) --
	(124.57,135.54) --
	(125.04,135.48) --
	(125.51,135.54) --
	(125.99,135.48) --
	(126.46,135.43) --
	(126.93,135.25) --
	(127.40,135.20) --
	(127.87,135.14) --
	(128.34,135.14) --
	(128.81,135.03) --
	(129.28,134.80) --
	(129.75,134.80) --
	(130.23,134.74) --
	(130.70,134.51) --
	(131.17,134.34) --
	(131.64,134.12) --
	(132.11,134.23) --
	(132.58,134.06);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,127.83) rectangle (137.69,197.70);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,197.70) rectangle (137.69,205.65);

\path[] ( 54.44,197.70) rectangle (137.69,205.65);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 69.67,199.40) {T=250, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,197.70) rectangle (137.69,205.65);
\end{scope}
\begin{scope}
\path[clip] (137.69,127.83) rectangle (220.94,197.70);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.45,156.69) --
	(218.97,156.69);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (145.16,151.90) --
	(145.63,152.76) --
	(146.10,151.40) --
	(146.57,152.26) --
	(147.05,152.63) --
	(147.52,153.00) --
	(147.99,152.63) --
	(148.46,154.23) --
	(148.93,154.60) --
	(149.40,155.33) --
	(149.87,152.76) --
	(150.34,152.88) --
	(150.81,152.51) --
	(151.29,152.39) --
	(151.76,152.39) --
	(152.23,151.90) --
	(152.70,152.02) --
	(153.17,151.28) --
	(153.64,149.32) --
	(154.11,148.95) --
	(154.58,149.68) --
	(155.05,148.46) --
	(155.53,148.09) --
	(156.00,148.33) --
	(156.47,147.47) --
	(156.94,148.09) --
	(157.41,147.97) --
	(157.88,146.86) --
	(158.35,145.75) --
	(158.82,145.75) --
	(159.29,145.75) --
	(159.77,146.00) --
	(160.24,146.00) --
	(160.71,145.75) --
	(161.18,144.65) --
	(161.65,144.28) --
	(162.12,143.91) --
	(162.59,142.81) --
	(163.06,142.19) --
	(163.53,142.07) --
	(164.01,142.07) --
	(164.48,142.07) --
	(164.95,142.44) --
	(165.42,141.82) --
	(165.89,141.58) --
	(166.36,142.19) --
	(166.83,140.84) --
	(167.30,141.70) --
	(167.77,141.46) --
	(168.25,140.23) --
	(168.72,140.35) --
	(169.19,140.47) --
	(169.66,140.47) --
	(170.13,140.84) --
	(170.60,140.10) --
	(171.07,139.98) --
	(171.54,140.10) --
	(172.02,140.35) --
	(172.49,139.49) --
	(172.96,138.39) --
	(173.43,138.02) --
	(173.90,137.53) --
	(174.37,137.77) --
	(174.84,137.40) --
	(175.31,136.91) --
	(175.78,136.54) --
	(176.26,136.91) --
	(176.73,136.79) --
	(177.20,136.91) --
	(177.67,137.40) --
	(178.14,137.40) --
	(178.61,137.65) --
	(179.08,137.28) --
	(179.55,137.16) --
	(180.02,137.40) --
	(180.50,137.53) --
	(180.97,136.79) --
	(181.44,136.30) --
	(181.91,136.05) --
	(182.38,136.30) --
	(182.85,136.05) --
	(183.32,135.68) --
	(183.79,135.32) --
	(184.26,135.44) --
	(184.74,135.44) --
	(185.21,135.19) --
	(185.68,135.56) --
	(186.15,135.32) --
	(186.62,135.68) --
	(187.09,135.68) --
	(187.56,135.81) --
	(188.03,135.68) --
	(188.50,135.44) --
	(188.98,135.68) --
	(189.45,135.32) --
	(189.92,135.07) --
	(190.39,134.82) --
	(190.86,134.70) --
	(191.33,134.82) --
	(191.80,134.82) --
	(192.27,134.33) --
	(192.74,134.95) --
	(193.22,134.58) --
	(193.69,134.33) --
	(194.16,134.58) --
	(194.63,134.58) --
	(195.10,134.33) --
	(195.57,134.09) --
	(196.04,134.33) --
	(196.51,134.33) --
	(196.99,133.96) --
	(197.46,133.47) --
	(197.93,133.23) --
	(198.40,133.23) --
	(198.87,133.47) --
	(199.34,133.35) --
	(199.81,133.23) --
	(200.28,133.35) --
	(200.75,133.47) --
	(201.23,132.98) --
	(201.70,132.86) --
	(202.17,132.86) --
	(202.64,132.86) --
	(203.11,132.86) --
	(203.58,132.86) --
	(204.05,132.86) --
	(204.52,132.61) --
	(204.99,132.49) --
	(205.47,132.74) --
	(205.94,132.61) --
	(206.41,132.37) --
	(206.88,132.37) --
	(207.35,132.24) --
	(207.82,132.37) --
	(208.29,132.37) --
	(208.76,132.49) --
	(209.23,132.37) --
	(209.71,132.37) --
	(210.18,132.37) --
	(210.65,132.24) --
	(211.12,132.37) --
	(211.59,132.24) --
	(212.06,132.24) --
	(212.53,132.24) --
	(213.00,132.24) --
	(213.47,132.24) --
	(213.95,132.24) --
	(214.42,132.24) --
	(214.89,132.12) --
	(215.36,132.12) --
	(215.83,132.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.69,127.83) rectangle (220.94,197.70);
\end{scope}
\begin{scope}
\path[clip] (137.69,197.70) rectangle (220.94,205.65);

\path[] (137.69,197.70) rectangle (220.94,205.65);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (147.24,199.40) {T=250, K=T/10, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.69,197.70) rectangle (220.94,205.65);
\end{scope}
\begin{scope}
\path[clip] (220.94,127.83) rectangle (304.19,197.70);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.70,156.69) --
	(302.22,156.69);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.41,162.09) --
	(228.88,160.49) --
	(229.35,160.62) --
	(229.82,160.00) --
	(230.29,161.11) --
	(230.77,160.00) --
	(231.24,158.53) --
	(231.71,159.14) --
	(232.18,159.88) --
	(232.65,157.79) --
	(233.12,157.91) --
	(233.59,159.14) --
	(234.06,158.77) --
	(234.53,157.79) --
	(235.01,158.53) --
	(235.48,159.39) --
	(235.95,159.26) --
	(236.42,158.53) --
	(236.89,158.65) --
	(237.36,159.39) --
	(237.83,159.88) --
	(238.30,159.14) --
	(238.77,157.30) --
	(239.25,156.93) --
	(239.72,156.44) --
	(240.19,156.56) --
	(240.66,157.18) --
	(241.13,157.05) --
	(241.60,158.16) --
	(242.07,157.18) --
	(242.54,157.55) --
	(243.01,156.56) --
	(243.49,157.05) --
	(243.96,157.05) --
	(244.43,156.32) --
	(244.90,156.56) --
	(245.37,155.83) --
	(245.84,156.93) --
	(246.31,157.30) --
	(246.78,157.30) --
	(247.26,156.44) --
	(247.73,156.81) --
	(248.20,157.67) --
	(248.67,157.30) --
	(249.14,157.30) --
	(249.61,156.93) --
	(250.08,156.81) --
	(250.55,156.69) --
	(251.02,157.05) --
	(251.50,156.56) --
	(251.97,156.32) --
	(252.44,155.83) --
	(252.91,156.69) --
	(253.38,156.07) --
	(253.85,156.32) --
	(254.32,156.07) --
	(254.79,156.19) --
	(255.26,156.44) --
	(255.74,155.83) --
	(256.21,154.72) --
	(256.68,155.09) --
	(257.15,155.09) --
	(257.62,155.83) --
	(258.09,155.95) --
	(258.56,154.84) --
	(259.03,155.09) --
	(259.50,156.19) --
	(259.98,155.46) --
	(260.45,155.46) --
	(260.92,153.61) --
	(261.39,153.74) --
	(261.86,154.47) --
	(262.33,154.11) --
	(262.80,153.86) --
	(263.27,154.11) --
	(263.74,154.23) --
	(264.22,154.47) --
	(264.69,154.72) --
	(265.16,153.74) --
	(265.63,154.84) --
	(266.10,154.23) --
	(266.57,154.84) --
	(267.04,154.11) --
	(267.51,153.74) --
	(267.98,153.74) --
	(268.46,154.47) --
	(268.93,155.21) --
	(269.40,154.47) --
	(269.87,154.35) --
	(270.34,154.11) --
	(270.81,153.37) --
	(271.28,152.76) --
	(271.75,153.12) --
	(272.23,152.88) --
	(272.70,152.88) --
	(273.17,153.00) --
	(273.64,152.51) --
	(274.11,152.88) --
	(274.58,151.90) --
	(275.05,151.90) --
	(275.52,151.28) --
	(275.99,151.53) --
	(276.47,151.40) --
	(276.94,151.16) --
	(277.41,151.16) --
	(277.88,151.04) --
	(278.35,150.67) --
	(278.82,151.40) --
	(279.29,150.79) --
	(279.76,150.91) --
	(280.23,150.79) --
	(280.71,150.30) --
	(281.18,150.18) --
	(281.65,150.42) --
	(282.12,149.81) --
	(282.59,149.81) --
	(283.06,149.44) --
	(283.53,149.32) --
	(284.00,149.19) --
	(284.47,148.70) --
	(284.95,148.70) --
	(285.42,148.95) --
	(285.89,149.07) --
	(286.36,149.44) --
	(286.83,149.68) --
	(287.30,149.07) --
	(287.77,149.19) --
	(288.24,149.32) --
	(288.71,148.58) --
	(289.19,148.83) --
	(289.66,149.32) --
	(290.13,148.09) --
	(290.60,148.09) --
	(291.07,147.97) --
	(291.54,148.09) --
	(292.01,148.09) --
	(292.48,147.47) --
	(292.96,148.09) --
	(293.43,147.60) --
	(293.90,147.60) --
	(294.37,147.35) --
	(294.84,147.47) --
	(295.31,146.86) --
	(295.78,146.86) --
	(296.25,146.98) --
	(296.72,145.63) --
	(297.20,145.51) --
	(297.67,145.26) --
	(298.14,145.02) --
	(298.61,145.14) --
	(299.08,146.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.94,127.83) rectangle (304.19,197.70);
\end{scope}
\begin{scope}
\path[clip] (220.94,197.70) rectangle (304.19,205.65);

\path[] (220.94,197.70) rectangle (304.19,205.65);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (236.17,199.40) {T=250, K=3, c=2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.94,197.70) rectangle (304.19,205.65);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,132.12) -- (393.13,132.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,144.40) -- (393.13,144.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,156.69) -- (393.13,156.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,168.97) -- (393.13,168.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,181.25) -- (393.13,181.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,193.53) -- (393.13,193.53);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (398.82,128.82) {0.00};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (398.82,141.10) {0.05};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (398.82,153.38) {0.10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (398.82,165.66) {0.15};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (398.82,177.94) {0.20};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (398.82,190.23) {0.25};
\end{scope}
\begin{scope}
\path[clip] (304.19,127.83) rectangle (387.44,197.70);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (306.95,156.69) --
	(385.47,156.69);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (311.66,167.62) --
	(312.13,167.25) --
	(312.60,169.34) --
	(313.07,171.91) --
	(313.54,170.69) --
	(314.01,171.67) --
	(314.49,171.18) --
	(314.96,172.41) --
	(315.43,173.02) --
	(315.90,173.76) --
	(316.37,173.76) --
	(316.84,175.23) --
	(317.31,175.11) --
	(317.78,176.34) --
	(318.25,176.21) --
	(318.73,176.09) --
	(319.20,176.09) --
	(319.67,177.07) --
	(320.14,177.44) --
	(320.61,178.06) --
	(321.08,178.18) --
	(321.55,176.83) --
	(322.02,176.95) --
	(322.50,178.55) --
	(322.97,179.90) --
	(323.44,176.70) --
	(323.91,178.67) --
	(324.38,179.16) --
	(324.85,178.67) --
	(325.32,178.42) --
	(325.79,178.67) --
	(326.26,179.53) --
	(326.74,180.76) --
	(327.21,182.11) --
	(327.68,181.00) --
	(328.15,182.23) --
	(328.62,182.23) --
	(329.09,182.11) --
	(329.56,181.86) --
	(330.03,181.86) --
	(330.50,181.25) --
	(330.98,181.62) --
	(331.45,181.13) --
	(331.92,181.13) --
	(332.39,181.86) --
	(332.86,181.13) --
	(333.33,181.74) --
	(333.80,180.51) --
	(334.27,179.16) --
	(334.74,179.41) --
	(335.22,179.90) --
	(335.69,181.62) --
	(336.16,180.39) --
	(336.63,179.90) --
	(337.10,180.51) --
	(337.57,180.27) --
	(338.04,178.92) --
	(338.51,178.30) --
	(338.98,179.41) --
	(339.46,181.13) --
	(339.93,180.27) --
	(340.40,179.78) --
	(340.87,179.04) --
	(341.34,180.39) --
	(341.81,180.27) --
	(342.28,180.51) --
	(342.75,180.02) --
	(343.22,178.55) --
	(343.70,177.69) --
	(344.17,177.56) --
	(344.64,177.07) --
	(345.11,177.20) --
	(345.58,178.42) --
	(346.05,177.44) --
	(346.52,176.95) --
	(346.99,177.32) --
	(347.47,176.58) --
	(347.94,177.32) --
	(348.41,177.20) --
	(348.88,176.46) --
	(349.35,175.85) --
	(349.82,176.09) --
	(350.29,176.95) --
	(350.76,175.72) --
	(351.23,175.97) --
	(351.71,176.95) --
	(352.18,175.85) --
	(352.65,176.09) --
	(353.12,175.97) --
	(353.59,175.85) --
	(354.06,176.34) --
	(354.53,176.83) --
	(355.00,176.70) --
	(355.47,176.09) --
	(355.95,175.11) --
	(356.42,175.35) --
	(356.89,174.13) --
	(357.36,174.74) --
	(357.83,174.25) --
	(358.30,172.65) --
	(358.77,174.00) --
	(359.24,173.27) --
	(359.71,172.65) --
	(360.19,171.91) --
	(360.66,171.55) --
	(361.13,172.04) --
	(361.60,172.53) --
	(362.07,171.30) --
	(362.54,169.95) --
	(363.01,169.83) --
	(363.48,169.83) --
	(363.95,168.97) --
	(364.43,170.32) --
	(364.90,170.07) --
	(365.37,169.70) --
	(365.84,169.58) --
	(366.31,169.95) --
	(366.78,169.83) --
	(367.25,169.70) --
	(367.72,168.60) --
	(368.19,167.74) --
	(368.67,166.51) --
	(369.14,166.63) --
	(369.61,166.76) --
	(370.08,166.02) --
	(370.55,165.53) --
	(371.02,165.04) --
	(371.49,164.42) --
	(371.96,163.81) --
	(372.44,164.30) --
	(372.91,164.55) --
	(373.38,164.55) --
	(373.85,163.69) --
	(374.32,163.07) --
	(374.79,161.97) --
	(375.26,162.21) --
	(375.73,161.11) --
	(376.20,161.11) --
	(376.68,160.37) --
	(377.15,160.25) --
	(377.62,160.37) --
	(378.09,159.39) --
	(378.56,159.88) --
	(379.03,159.39) --
	(379.50,157.42) --
	(379.97,157.30) --
	(380.44,157.42) --
	(380.92,155.46) --
	(381.39,154.35) --
	(381.86,154.60) --
	(382.33,153.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.19,127.83) rectangle (387.44,197.70);
\end{scope}
\begin{scope}
\path[clip] (304.19,197.70) rectangle (387.44,205.65);

\path[] (304.19,197.70) rectangle (387.44,205.65);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (313.74,199.40) {T=250, K=T/10, c=2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.19,197.70) rectangle (387.44,205.65);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.20,283.46) -- ( 57.20,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 80.76,283.46) -- ( 80.76,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (104.31,283.46) -- (104.31,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (127.87,283.46) -- (127.87,289.15);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,209.94) -- ( 48.75,209.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,222.22) -- ( 48.75,222.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,234.50) -- ( 48.75,234.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,246.78) -- ( 48.75,246.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,259.07) -- ( 48.75,259.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,271.35) -- ( 48.75,271.35);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,206.63) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,218.91) {0.05};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,231.20) {0.10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,243.48) {0.15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,255.76) {0.20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,268.04) {0.25};
\end{scope}
\begin{scope}
\path[clip] ( 54.44,205.65) rectangle (137.69,275.51);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 57.20,234.50) --
	(135.72,234.50);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.98,233.38) --
	( 70.16,233.69) --
	( 71.33,232.64) --
	( 72.51,232.40) --
	( 73.69,231.72) --
	( 74.87,231.17) --
	( 76.05,230.37) --
	( 77.22,229.51) --
	( 78.40,229.14) --
	( 79.58,228.72) --
	( 80.76,228.53) --
	( 81.93,228.10) --
	( 83.11,228.22) --
	( 84.29,227.49) --
	( 85.47,226.81) --
	( 86.65,226.20) --
	( 87.82,225.40) --
	( 89.00,225.46) --
	( 90.18,225.03) --
	( 91.36,224.60) --
	( 92.53,223.99) --
	( 93.71,223.32) --
	( 94.89,222.82) --
	( 96.07,222.33) --
	( 97.25,222.70) --
	( 98.42,222.09) --
	( 99.60,221.72) --
	(100.78,220.92) --
	(101.96,220.55) --
	(103.14,220.12) --
	(104.31,220.25) --
	(105.49,219.51) --
	(106.67,219.20) --
	(107.85,218.65) --
	(109.02,217.73) --
	(110.20,218.10) --
	(111.38,217.42) --
	(112.56,217.06) --
	(113.74,216.69) --
	(114.91,216.81) --
	(116.09,216.81) --
	(117.27,216.38) --
	(118.45,216.01) --
	(119.63,215.64) --
	(120.80,215.28) --
	(121.98,215.22) --
	(123.16,215.03) --
	(124.34,214.79) --
	(125.51,214.23) --
	(126.69,213.93) --
	(127.87,213.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,205.65) rectangle (137.69,275.51);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,275.51) rectangle (137.69,283.46);

\path[] ( 54.44,275.51) rectangle (137.69,283.46);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 69.67,277.22) {T=100, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,275.51) rectangle (137.69,283.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.45,283.46) -- (140.45,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (164.01,283.46) -- (164.01,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (187.56,283.46) -- (187.56,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.12,283.46) -- (211.12,289.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (140.45,294.85) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (164.01,294.85) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (187.56,294.85) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (211.12,294.85) {0.6};
\end{scope}
\begin{scope}
\path[clip] (137.69,205.65) rectangle (220.94,275.51);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (140.45,234.50) --
	(218.97,234.50);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.23,230.78) --
	(153.41,229.22) --
	(154.58,228.02) --
	(155.76,227.06) --
	(156.94,227.54) --
	(158.12,225.99) --
	(159.29,225.03) --
	(160.47,224.19) --
	(161.65,224.79) --
	(162.83,224.07) --
	(164.01,223.11) --
	(165.18,222.51) --
	(166.36,221.32) --
	(167.54,221.56) --
	(168.72,221.08) --
	(169.90,220.96) --
	(171.07,220.84) --
	(172.25,220.12) --
	(173.43,219.88) --
	(174.61,219.88) --
	(175.78,219.40) --
	(176.96,218.92) --
	(178.14,217.36) --
	(179.32,217.60) --
	(180.50,217.36) --
	(181.67,217.00) --
	(182.85,216.41) --
	(184.03,215.45) --
	(185.21,215.45) --
	(186.38,214.85) --
	(187.56,215.45) --
	(188.74,215.21) --
	(189.92,214.85) --
	(191.10,214.73) --
	(192.27,214.13) --
	(193.45,214.25) --
	(194.63,214.61) --
	(195.81,213.77) --
	(196.99,213.65) --
	(198.16,213.65) --
	(199.34,213.41) --
	(200.52,212.93) --
	(201.70,212.57) --
	(202.87,212.45) --
	(204.05,212.45) --
	(205.23,211.85) --
	(206.41,211.97) --
	(207.59,211.38) --
	(208.76,211.49) --
	(209.94,211.38) --
	(211.12,211.26);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.69,205.65) rectangle (220.94,275.51);
\end{scope}
\begin{scope}
\path[clip] (137.69,275.51) rectangle (220.94,283.46);

\path[] (137.69,275.51) rectangle (220.94,283.46);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (147.24,277.22) {T=100, K=T/10, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (137.69,275.51) rectangle (220.94,283.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.70,283.46) -- (223.70,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.26,283.46) -- (247.26,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (270.81,283.46) -- (270.81,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (294.37,283.46) -- (294.37,289.15);
\end{scope}
\begin{scope}
\path[clip] (220.94,205.65) rectangle (304.19,275.51);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (223.70,234.50) --
	(302.22,234.50);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (235.48,244.20) --
	(236.65,239.29) --
	(237.83,237.69) --
	(239.01,238.43) --
	(240.19,237.57) --
	(241.37,237.94) --
	(242.54,238.92) --
	(243.72,235.98) --
	(244.90,234.62) --
	(246.08,234.62) --
	(247.26,235.85) --
	(248.43,234.38) --
	(249.61,234.99) --
	(250.79,235.61) --
	(251.97,235.48) --
	(253.14,236.10) --
	(254.32,234.01) --
	(255.50,234.13) --
	(256.68,233.89) --
	(257.86,233.27) --
	(259.03,233.15) --
	(260.21,232.05) --
	(261.39,232.41) --
	(262.57,232.90) --
	(263.74,231.31) --
	(264.92,231.31) --
	(266.10,231.43) --
	(267.28,230.94) --
	(268.46,230.33) --
	(269.63,231.68) --
	(270.81,231.06) --
	(271.99,231.31) --
	(273.17,231.06) --
	(274.35,230.08) --
	(275.52,229.34) --
	(276.70,229.71) --
	(277.88,228.36) --
	(279.06,228.73) --
	(280.23,227.75) --
	(281.41,228.61) --
	(282.59,226.89) --
	(283.77,226.76) --
	(284.95,226.40) --
	(286.12,225.78) --
	(287.30,225.17) --
	(288.48,225.41) --
	(289.66,225.04) --
	(290.83,224.68) --
	(292.01,224.43) --
	(293.19,224.31) --
	(294.37,223.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.94,205.65) rectangle (304.19,275.51);
\end{scope}
\begin{scope}
\path[clip] (220.94,275.51) rectangle (304.19,283.46);

\path[] (220.94,275.51) rectangle (304.19,283.46);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (236.17,277.22) {T=100, K=3, c=2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (220.94,275.51) rectangle (304.19,283.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (306.95,283.46) -- (306.95,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.50,283.46) -- (330.50,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (354.06,283.46) -- (354.06,289.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.62,283.46) -- (377.62,289.15);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (306.95,294.85) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (330.50,294.85) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (354.06,294.85) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (377.62,294.85) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,209.94) -- (393.13,209.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,222.22) -- (393.13,222.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,234.50) -- (393.13,234.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,246.78) -- (393.13,246.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,259.07) -- (393.13,259.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.44,271.35) -- (393.13,271.35);
\end{scope}
\begin{scope}
\path[clip] (304.19,205.65) rectangle (387.44,275.51);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (306.95,234.50) --
	(385.47,234.50);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.73,245.19) --
	(319.90,243.84) --
	(321.08,246.54) --
	(322.26,247.27) --
	(323.44,246.78) --
	(324.62,247.15) --
	(325.79,246.66) --
	(326.97,246.42) --
	(328.15,245.56) --
	(329.33,245.92) --
	(330.50,246.91) --
	(331.68,245.80) --
	(332.86,244.70) --
	(334.04,245.06) --
	(335.22,246.05) --
	(336.39,246.17) --
	(337.57,244.94) --
	(338.75,244.45) --
	(339.93,246.17) --
	(341.10,244.45) --
	(342.28,243.47) --
	(343.46,243.34) --
	(344.64,241.75) --
	(345.82,241.50) --
	(346.99,241.38) --
	(348.17,242.48) --
	(349.35,241.87) --
	(350.53,240.77) --
	(351.71,240.77) --
	(352.88,240.40) --
	(354.06,239.91) --
	(355.24,239.29) --
	(356.42,237.82) --
	(357.59,238.68) --
	(358.77,238.31) --
	(359.95,237.94) --
	(361.13,239.05) --
	(362.31,238.06) --
	(363.48,236.47) --
	(364.66,235.98) --
	(365.84,236.47) --
	(367.02,234.62) --
	(368.19,233.64) --
	(369.37,232.05) --
	(370.55,231.68) --
	(371.73,232.05) --
	(372.91,231.92) --
	(374.08,232.05) --
	(375.26,231.92) --
	(376.44,230.20) --
	(377.62,229.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.19,205.65) rectangle (387.44,275.51);
\end{scope}
\begin{scope}
\path[clip] (304.19,275.51) rectangle (387.44,283.46);

\path[] (304.19,275.51) rectangle (387.44,283.46);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (313.74,277.22) {T=100, K=T/10, c=2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,325.21);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.19,275.51) rectangle (387.44,283.46);
\end{scope}
\end{tikzpicture}