### 2017-04-21 floats/mc-ftest.tex

\begin{table}[tb]
 % latex table generated in R 3.0.2 by xtable 1.7-0 package
% Thu Mar 27 13:12:32 2014
\begin{tabularx}{\textwidth}{llCCC}
  \toprule &&&
 \multicolumn{2}{c}{Conditional rejection probability (size)} \\
 \cmidrule(l){4-5} $K_1$ & $K_2$ & $T$ & $| \theta_2 |_2 = 0$ & $| \theta_2 |_2 = 1$ \\ 
  \midrule $2$ & $3$ & $100$ & $0.104$ & $0.485$ \\ 
   &  & $250$ & $0.102$ & $0.528$ \\ 
   &  & $500$ & $0.096$ & $0.503$ \\ 
  $T/50$ & $T/10$ & $100$ & $0.106$ & $0.701$ \\ 
   &  & $250$ & $0.108$ & $0.904$ \\ 
   &  & $500$ & $0.104$ & $0.989$ \\ 
   \bottomrule \end{tabularx}
 \caption{Simulated rejection probabilities for the $F$-test under
the null hypothesis that the benchmark model will forecast better,
$\E_T \oosB \leq 0$. Nominal size is 10\% and values greater than
10\% indicate that the test rejects the benchmark model too often.
See Section~\ref{sec:simulation-design} for a discussion of the
simulation design.}
\label{fig:ftest}
\end{table}