### 2017-04-21 floats/mc-mccrackensize-rec.tex

% Created by tikzDevice version 0.6.2-92-0ad2792 on 2014-03-27 13:56:00
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (361.35,397.48);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);

\path[] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (184.81, 12.04) {$P/T$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 14.29,198.90) {Rejection Probability};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 57.82) -- ( 48.75, 57.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 72.33) -- ( 48.75, 72.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 86.84) -- ( 48.75, 86.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,101.35) -- ( 48.75,101.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,115.86) -- ( 48.75,115.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,130.37) -- ( 48.75,130.37);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 54.51) {0.10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 69.02) {0.15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 83.53) {0.20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 98.04) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,112.55) {0.30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,127.06) {0.35};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76, 50.02) -- ( 58.76, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65, 50.02) -- ( 95.65, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (132.54, 50.02) -- (132.54, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.42, 50.02) -- (169.42, 44.32);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 58.76, 32.02) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 95.65, 32.02) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (132.54, 32.02) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (169.42, 32.02) {0.6};
\end{scope}
\begin{scope}
\path[clip] ( 54.44, 50.02) rectangle (184.81,143.97);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76, 57.82) --
	(181.72, 57.82);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.45, 59.12) --
	( 62.82, 60.14) --
	( 63.19, 60.86) --
	( 63.56, 60.14) --
	( 63.92, 59.70) --
	( 64.29, 61.88) --
	( 64.66, 62.46) --
	( 65.03, 61.30) --
	( 65.40, 60.14) --
	( 65.77, 60.57) --
	( 66.14, 60.86) --
	( 66.51, 60.57) --
	( 66.88, 60.86) --
	( 67.24, 60.43) --
	( 67.61, 60.28) --
	( 67.98, 61.01) --
	( 68.35, 60.28) --
	( 68.72, 59.56) --
	( 69.09, 61.88) --
	( 69.46, 62.46) --
	( 69.83, 61.30) --
	( 70.20, 63.04) --
	( 70.56, 62.75) --
	( 70.93, 63.33) --
	( 71.30, 63.18) --
	( 71.67, 64.64) --
	( 72.04, 64.06) --
	( 72.41, 64.20) --
	( 72.78, 64.78) --
	( 73.15, 64.64) --
	( 73.52, 64.93) --
	( 73.88, 65.51) --
	( 74.25, 65.36) --
	( 74.62, 65.94) --
	( 74.99, 65.94) --
	( 75.36, 65.36) --
	( 75.73, 65.07) --
	( 76.10, 63.77) --
	( 76.47, 63.77) --
	( 76.84, 64.35) --
	( 77.20, 64.49) --
	( 77.57, 63.62) --
	( 77.94, 62.75) --
	( 78.31, 62.89) --
	( 78.68, 63.62) --
	( 79.05, 63.77) --
	( 79.42, 64.06) --
	( 79.79, 64.49) --
	( 80.16, 63.48) --
	( 80.52, 62.46) --
	( 80.89, 62.75) --
	( 81.26, 62.17) --
	( 81.63, 61.30) --
	( 82.00, 61.44) --
	( 82.37, 61.44) --
	( 82.74, 61.30) --
	( 83.11, 61.59) --
	( 83.48, 61.88) --
	( 83.84, 62.31) --
	( 84.21, 62.31) --
	( 84.58, 62.02) --
	( 84.95, 63.04) --
	( 85.32, 64.49) --
	( 85.69, 63.18) --
	( 86.06, 62.75) --
	( 86.43, 63.91) --
	( 86.79, 63.33) --
	( 87.16, 63.62) --
	( 87.53, 63.48) --
	( 87.90, 63.48) --
	( 88.27, 63.18) --
	( 88.64, 63.04) --
	( 89.01, 64.20) --
	( 89.38, 64.20) --
	( 89.75, 64.78) --
	( 90.11, 64.20) --
	( 90.48, 63.33) --
	( 90.85, 63.04) --
	( 91.22, 62.60) --
	( 91.59, 63.62) --
	( 91.96, 63.04) --
	( 92.33, 62.89) --
	( 92.70, 62.75) --
	( 93.07, 62.89) --
	( 93.43, 63.04) --
	( 93.80, 62.31) --
	( 94.17, 62.46) --
	( 94.54, 62.17) --
	( 94.91, 62.31) --
	( 95.28, 62.46) --
	( 95.65, 63.04) --
	( 96.02, 64.35) --
	( 96.39, 63.91) --
	( 96.75, 64.20) --
	( 97.12, 62.31) --
	( 97.49, 62.17) --
	( 97.86, 61.73) --
	( 98.23, 62.31) --
	( 98.60, 62.46) --
	( 98.97, 61.44) --
	( 99.34, 62.02) --
	( 99.71, 61.88) --
	(100.07, 62.31) --
	(100.44, 61.73) --
	(100.81, 61.44) --
	(101.18, 60.57) --
	(101.55, 60.86) --
	(101.92, 60.28) --
	(102.29, 60.28) --
	(102.66, 60.72) --
	(103.03, 61.44) --
	(103.39, 60.72) --
	(103.76, 60.72) --
	(104.13, 61.15) --
	(104.50, 61.44) --
	(104.87, 60.28) --
	(105.24, 59.99) --
	(105.61, 59.99) --
	(105.98, 60.86) --
	(106.35, 60.28) --
	(106.71, 61.01) --
	(107.08, 60.86) --
	(107.45, 59.70) --
	(107.82, 60.14) --
	(108.19, 60.14) --
	(108.56, 59.41) --
	(108.93, 59.12) --
	(109.30, 58.54) --
	(109.67, 59.12) --
	(110.03, 59.85) --
	(110.40, 59.70) --
	(110.77, 59.27) --
	(111.14, 59.41) --
	(111.51, 58.98) --
	(111.88, 58.54) --
	(112.25, 58.40) --
	(112.62, 58.54) --
	(112.99, 58.25) --
	(113.35, 59.70) --
	(113.72, 59.70) --
	(114.09, 59.85) --
	(114.46, 58.98) --
	(114.83, 58.83) --
	(115.20, 59.41) --
	(115.57, 58.69) --
	(115.94, 58.83) --
	(116.31, 58.54) --
	(116.67, 57.67) --
	(117.04, 58.11) --
	(117.41, 57.53) --
	(117.78, 58.40) --
	(118.15, 57.96) --
	(118.52, 58.25) --
	(118.89, 57.96) --
	(119.26, 58.25) --
	(119.63, 58.54) --
	(119.99, 59.70) --
	(120.36, 59.12) --
	(120.73, 59.41) --
	(121.10, 59.85) --
	(121.47, 59.56) --
	(121.84, 59.70) --
	(122.21, 58.83) --
	(122.58, 58.98) --
	(122.95, 59.12) --
	(123.31, 58.54) --
	(123.68, 59.27) --
	(124.05, 59.27) --
	(124.42, 58.69) --
	(124.79, 58.69) --
	(125.16, 58.40) --
	(125.53, 58.54) --
	(125.90, 57.82) --
	(126.27, 58.40) --
	(126.63, 58.69) --
	(127.00, 59.27) --
	(127.37, 58.83) --
	(127.74, 59.27) --
	(128.11, 59.41) --
	(128.48, 59.56) --
	(128.85, 59.12) --
	(129.22, 59.12) --
	(129.59, 58.25) --
	(129.95, 58.40) --
	(130.32, 58.69) --
	(130.69, 58.40) --
	(131.06, 58.83) --
	(131.43, 58.69) --
	(131.80, 58.83) --
	(132.17, 58.83) --
	(132.54, 59.27) --
	(132.91, 59.27) --
	(133.27, 58.69) --
	(133.64, 58.69) --
	(134.01, 59.41) --
	(134.38, 58.98) --
	(134.75, 58.83) --
	(135.12, 59.70) --
	(135.49, 59.27) --
	(135.86, 59.12) --
	(136.23, 58.54) --
	(136.59, 59.56) --
	(136.96, 58.98) --
	(137.33, 58.98) --
	(137.70, 59.56) --
	(138.07, 59.27) --
	(138.44, 59.56) --
	(138.81, 58.69) --
	(139.18, 59.27) --
	(139.55, 59.12) --
	(139.91, 58.83) --
	(140.28, 58.11) --
	(140.65, 57.82) --
	(141.02, 58.54) --
	(141.39, 58.69) --
	(141.76, 58.40) --
	(142.13, 57.82) --
	(142.50, 58.40) --
	(142.87, 58.11) --
	(143.23, 57.82) --
	(143.60, 57.53) --
	(143.97, 58.40) --
	(144.34, 58.54) --
	(144.71, 58.54) --
	(145.08, 58.54) --
	(145.45, 58.83) --
	(145.82, 57.96) --
	(146.19, 58.54) --
	(146.55, 58.40) --
	(146.92, 58.54) --
	(147.29, 58.69) --
	(147.66, 58.40) --
	(148.03, 59.12) --
	(148.40, 58.25) --
	(148.77, 58.40) --
	(149.14, 57.67) --
	(149.51, 57.96) --
	(149.87, 57.67) --
	(150.24, 56.51) --
	(150.61, 57.82) --
	(150.98, 57.96) --
	(151.35, 58.11) --
	(151.72, 57.67) --
	(152.09, 57.96) --
	(152.46, 58.11) --
	(152.83, 58.25) --
	(153.19, 57.53) --
	(153.56, 57.82) --
	(153.93, 57.24) --
	(154.30, 57.82) --
	(154.67, 57.82) --
	(155.04, 57.53) --
	(155.41, 56.51) --
	(155.78, 56.95) --
	(156.15, 57.96) --
	(156.51, 58.11) --
	(156.88, 58.11) --
	(157.25, 58.25) --
	(157.62, 58.40) --
	(157.99, 58.25) --
	(158.36, 58.25) --
	(158.73, 58.83) --
	(159.10, 57.96) --
	(159.47, 57.38) --
	(159.83, 57.24) --
	(160.20, 56.66) --
	(160.57, 56.80) --
	(160.94, 56.22) --
	(161.31, 56.66) --
	(161.68, 57.09) --
	(162.05, 56.95) --
	(162.42, 56.51) --
	(162.78, 56.66) --
	(163.15, 55.78) --
	(163.52, 55.93) --
	(163.89, 55.78) --
	(164.26, 56.22) --
	(164.63, 56.51) --
	(165.00, 56.66) --
	(165.37, 57.38) --
	(165.74, 57.24) --
	(166.10, 56.66) --
	(166.47, 56.80) --
	(166.84, 56.80) --
	(167.21, 56.80) --
	(167.58, 56.66) --
	(167.95, 56.37) --
	(168.32, 56.22) --
	(168.69, 56.07) --
	(169.06, 57.24) --
	(169.42, 56.66) --
	(169.79, 57.67) --
	(170.16, 57.53) --
	(170.53, 58.54) --
	(170.90, 58.25) --
	(171.27, 57.67) --
	(171.64, 58.40) --
	(172.01, 58.25) --
	(172.38, 58.54) --
	(172.74, 57.82) --
	(173.11, 57.96) --
	(173.48, 58.54) --
	(173.85, 58.54) --
	(174.22, 57.82) --
	(174.59, 57.96) --
	(174.96, 57.67) --
	(175.33, 58.25) --
	(175.70, 57.96) --
	(176.06, 57.82) --
	(176.43, 57.53) --
	(176.80, 56.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 50.02) rectangle (184.81,143.97);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,143.97) rectangle (184.81,151.92);

\path[] ( 54.44,143.97) rectangle (184.81,151.92);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 93.23,145.67) {T=500, K=3, c=0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,143.97) rectangle (184.81,151.92);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12, 50.02) -- (189.12, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.01, 50.02) -- (226.01, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.90, 50.02) -- (262.90, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.79, 50.02) -- (299.79, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17, 57.82) -- (320.86, 57.82);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17, 72.33) -- (320.86, 72.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17, 86.84) -- (320.86, 86.84);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,101.35) -- (320.86,101.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,115.86) -- (320.86,115.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,130.37) -- (320.86,130.37);
\end{scope}
\begin{scope}
\path[clip] (184.81, 50.02) rectangle (315.17,143.97);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12, 57.82) --
	(312.08, 57.82);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.81, 64.06) --
	(193.18, 64.78) --
	(193.55, 64.49) --
	(193.92, 65.51) --
	(194.29, 64.06) --
	(194.66, 65.51) --
	(195.03, 64.35) --
	(195.39, 67.10) --
	(195.76, 66.96) --
	(196.13, 66.67) --
	(196.50, 65.22) --
	(196.87, 66.67) --
	(197.24, 67.54) --
	(197.61, 66.23) --
	(197.98, 66.09) --
	(198.35, 68.12) --
	(198.71, 67.10) --
	(199.08, 68.70) --
	(199.45, 68.55) --
	(199.82, 68.70) --
	(200.19, 69.28) --
	(200.56, 71.17) --
	(200.93, 72.76) --
	(201.30, 73.34) --
	(201.67, 73.05) --
	(202.03, 74.36) --
	(202.40, 74.79) --
	(202.77, 76.10) --
	(203.14, 77.11) --
	(203.51, 77.26) --
	(203.88, 78.13) --
	(204.25, 77.26) --
	(204.62, 79.15) --
	(204.99, 78.57) --
	(205.35, 80.02) --
	(205.72, 79.87) --
	(206.09, 81.18) --
	(206.46, 81.90) --
	(206.83, 81.03) --
	(207.20, 81.03) --
	(207.57, 80.89) --
	(207.94, 81.03) --
	(208.31, 81.61) --
	(208.67, 81.90) --
	(209.04, 81.61) --
	(209.41, 81.18) --
	(209.78, 82.48) --
	(210.15, 81.76) --
	(210.52, 82.19) --
	(210.89, 82.77) --
	(211.26, 82.19) --
	(211.63, 82.48) --
	(211.99, 82.19) --
	(212.36, 82.05) --
	(212.73, 82.34) --
	(213.10, 82.77) --
	(213.47, 83.50) --
	(213.84, 84.66) --
	(214.21, 84.66) --
	(214.58, 85.68) --
	(214.95, 85.39) --
	(215.31, 86.55) --
	(215.68, 86.11) --
	(216.05, 86.26) --
	(216.42, 85.24) --
	(216.79, 86.26) --
	(217.16, 86.84) --
	(217.53, 87.56) --
	(217.90, 86.40) --
	(218.27, 88.14) --
	(218.63, 88.72) --
	(219.00, 88.43) --
	(219.37, 89.45) --
	(219.74, 89.30) --
	(220.11, 89.45) --
	(220.48, 90.75) --
	(220.85, 90.90) --
	(221.22, 91.04) --
	(221.58, 90.90) --
	(221.95, 90.32) --
	(222.32, 91.48) --
	(222.69, 91.48) --
	(223.06, 91.19) --
	(223.43, 90.03) --
	(223.80, 89.45) --
	(224.17, 90.03) --
	(224.54, 90.46) --
	(224.90, 89.74) --
	(225.27, 90.61) --
	(225.64, 90.61) --
	(226.01, 90.46) --
	(226.38, 90.61) --
	(226.75, 92.35) --
	(227.12, 91.62) --
	(227.49, 91.33) --
	(227.86, 91.33) --
	(228.22, 91.04) --
	(228.59, 91.92) --
	(228.96, 92.35) --
	(229.33, 92.06) --
	(229.70, 91.77) --
	(230.07, 91.92) --
	(230.44, 92.79) --
	(230.81, 92.79) --
	(231.18, 93.22) --
	(231.54, 93.22) --
	(231.91, 93.22) --
	(232.28, 93.80) --
	(232.65, 94.09) --
	(233.02, 94.67) --
	(233.39, 95.11) --
	(233.76, 94.96) --
	(234.13, 94.38) --
	(234.50, 95.11) --
	(234.86, 95.54) --
	(235.23, 94.96) --
	(235.60, 95.69) --
	(235.97, 95.54) --
	(236.34, 96.56) --
	(236.71, 95.54) --
	(237.08, 95.54) --
	(237.45, 96.12) --
	(237.82, 97.43) --
	(238.18, 96.56) --
	(238.55, 95.98) --
	(238.92, 96.56) --
	(239.29, 96.85) --
	(239.66, 96.85) --
	(240.03, 97.14) --
	(240.40, 97.43) --
	(240.77, 97.14) --
	(241.14, 97.43) --
	(241.50, 97.57) --
	(241.87, 97.14) --
	(242.24, 97.28) --
	(242.61, 98.15) --
	(242.98, 98.74) --
	(243.35, 98.44) --
	(243.72, 99.61) --
	(244.09, 99.46) --
	(244.46, 99.61) --
	(244.82,100.48) --
	(245.19,100.48) --
	(245.56,100.19) --
	(245.93,101.06) --
	(246.30,101.64) --
	(246.67,101.49) --
	(247.04,102.36) --
	(247.41,101.49) --
	(247.78,101.78) --
	(248.14,102.65) --
	(248.51,103.09) --
	(248.88,103.23) --
	(249.25,103.96) --
	(249.62,103.67) --
	(249.99,103.09) --
	(250.36,103.38) --
	(250.73,103.52) --
	(251.10,103.67) --
	(251.46,103.09) --
	(251.83,103.96) --
	(252.20,103.09) --
	(252.57,104.10) --
	(252.94,103.52) --
	(253.31,103.09) --
	(253.68,103.23) --
	(254.05,102.94) --
	(254.42,102.94) --
	(254.78,104.25) --
	(255.15,104.10) --
	(255.52,104.83) --
	(255.89,103.81) --
	(256.26,106.28) --
	(256.63,105.70) --
	(257.00,105.41) --
	(257.37,105.85) --
	(257.74,106.43) --
	(258.10,106.43) --
	(258.47,105.26) --
	(258.84,105.99) --
	(259.21,105.85) --
	(259.58,106.86) --
	(259.95,106.72) --
	(260.32,107.30) --
	(260.69,107.15) --
	(261.06,106.72) --
	(261.42,108.02) --
	(261.79,107.44) --
	(262.16,107.15) --
	(262.53,106.86) --
	(262.90,107.01) --
	(263.27,108.75) --
	(263.64,109.04) --
	(264.01,109.18) --
	(264.38,110.05) --
	(264.74,109.91) --
	(265.11,109.76) --
	(265.48,110.49) --
	(265.85,111.07) --
	(266.22,111.07) --
	(266.59,111.36) --
	(266.96,111.65) --
	(267.33,110.34) --
	(267.70,110.34) --
	(268.06,110.49) --
	(268.43,110.92) --
	(268.80,110.78) --
	(269.17,111.79) --
	(269.54,111.94) --
	(269.91,113.25) --
	(270.28,113.25) --
	(270.65,113.10) --
	(271.02,113.83) --
	(271.38,114.41) --
	(271.75,114.99) --
	(272.12,114.12) --
	(272.49,113.68) --
	(272.86,114.70) --
	(273.23,114.84) --
	(273.60,114.12) --
	(273.97,114.55) --
	(274.34,115.57) --
	(274.70,114.99) --
	(275.07,116.00) --
	(275.44,116.29) --
	(275.81,115.71) --
	(276.18,116.44) --
	(276.55,117.16) --
	(276.92,117.74) --
	(277.29,118.90) --
	(277.66,118.90) --
	(278.02,118.76) --
	(278.39,119.48) --
	(278.76,119.48) --
	(279.13,118.76) --
	(279.50,120.21) --
	(279.87,120.50) --
	(280.24,120.21) --
	(280.61,120.94) --
	(280.98,121.23) --
	(281.34,121.37) --
	(281.71,121.95) --
	(282.08,120.94) --
	(282.45,121.23) --
	(282.82,120.94) --
	(283.19,122.24) --
	(283.56,121.95) --
	(283.93,122.39) --
	(284.30,121.66) --
	(284.66,121.66) --
	(285.03,121.08) --
	(285.40,122.39) --
	(285.77,122.24) --
	(286.14,122.68) --
	(286.51,121.95) --
	(286.88,122.97) --
	(287.25,122.68) --
	(287.62,122.24) --
	(287.98,123.26) --
	(288.35,123.11) --
	(288.72,123.84) --
	(289.09,123.98) --
	(289.46,124.42) --
	(289.83,124.56) --
	(290.20,125.00) --
	(290.57,125.29) --
	(290.94,125.29) --
	(291.30,125.29) --
	(291.67,125.43) --
	(292.04,125.00) --
	(292.41,124.85) --
	(292.78,123.40) --
	(293.15,123.98) --
	(293.52,124.27) --
	(293.89,124.42) --
	(294.26,125.00) --
	(294.62,125.14) --
	(294.99,124.71) --
	(295.36,125.72) --
	(295.73,124.85) --
	(296.10,124.85) --
	(296.47,125.29) --
	(296.84,126.45) --
	(297.21,126.01) --
	(297.58,125.43) --
	(297.94,125.72) --
	(298.31,127.03) --
	(298.68,127.47) --
	(299.05,126.45) --
	(299.42,126.74) --
	(299.79,126.59) --
	(300.16,126.16) --
	(300.53,126.74) --
	(300.89,126.88) --
	(301.26,127.18) --
	(301.63,127.61) --
	(302.00,128.34) --
	(302.37,127.76) --
	(302.74,129.06) --
	(303.11,129.50) --
	(303.48,128.19) --
	(303.85,130.51) --
	(304.21,130.66) --
	(304.58,129.93) --
	(304.95,131.09) --
	(305.32,131.38) --
	(305.69,131.09) --
	(306.06,132.69) --
	(306.43,132.25) --
	(306.80,131.24) --
	(307.17,131.53);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81, 50.02) rectangle (315.17,143.97);
\end{scope}
\begin{scope}
\path[clip] (184.81,143.97) rectangle (315.17,151.92);

\path[] (184.81,143.97) rectangle (315.17,151.92);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (223.60,145.67) {T=500, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,143.97) rectangle (315.17,151.92);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,159.72) -- ( 48.75,159.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,174.23) -- ( 48.75,174.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,188.74) -- ( 48.75,188.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,203.25) -- ( 48.75,203.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,217.76) -- ( 48.75,217.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,232.27) -- ( 48.75,232.27);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,151.92) rectangle (184.81,245.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76,159.72) --
	(181.72,159.72);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.14,163.79) --
	( 66.88,164.37) --
	( 67.61,163.64) --
	( 68.35,163.93) --
	( 69.09,163.64) --
	( 69.83,166.25) --
	( 70.56,167.85) --
	( 71.30,168.14) --
	( 72.04,170.90) --
	( 72.78,169.59) --
	( 73.52,169.30) --
	( 74.25,169.01) --
	( 74.99,168.14) --
	( 75.73,169.44) --
	( 76.47,170.46) --
	( 77.20,170.17) --
	( 77.94,170.46) --
	( 78.68,170.31) --
	( 79.42,169.30) --
	( 80.16,168.86) --
	( 80.89,167.41) --
	( 81.63,166.40) --
	( 82.37,166.69) --
	( 83.11,167.27) --
	( 83.84,166.25) --
	( 84.58,165.53) --
	( 85.32,166.54) --
	( 86.06,165.53) --
	( 86.79,166.25) --
	( 87.53,166.69) --
	( 88.27,167.41) --
	( 89.01,167.12) --
	( 89.75,169.01) --
	( 90.48,168.14) --
	( 91.22,168.14) --
	( 91.96,168.57) --
	( 92.70,166.40) --
	( 93.43,165.96) --
	( 94.17,165.53) --
	( 94.91,166.11) --
	( 95.65,166.11) --
	( 96.39,165.24) --
	( 97.12,165.53) --
	( 97.86,165.24) --
	( 98.60,165.38) --
	( 99.34,165.96) --
	(100.07,164.66) --
	(100.81,164.80) --
	(101.55,165.09) --
	(102.29,165.09) --
	(103.03,164.22) --
	(103.76,163.79) --
	(104.50,163.50) --
	(105.24,162.62) --
	(105.98,163.06) --
	(106.71,162.19) --
	(107.45,161.90) --
	(108.19,161.61) --
	(108.93,162.04) --
	(109.67,161.03) --
	(110.40,161.75) --
	(111.14,160.74) --
	(111.88,161.32) --
	(112.62,161.90) --
	(113.35,161.46) --
	(114.09,161.17) --
	(114.83,161.61) --
	(115.57,161.61) --
	(116.31,162.19) --
	(117.04,161.32) --
	(117.78,161.46) --
	(118.52,162.19) --
	(119.26,162.19) --
	(119.99,163.06) --
	(120.73,162.48) --
	(121.47,163.35) --
	(122.21,163.35) --
	(122.95,162.48) --
	(123.68,162.19) --
	(124.42,162.48) --
	(125.16,161.90) --
	(125.90,162.04) --
	(126.63,162.48) --
	(127.37,162.91) --
	(128.11,162.77) --
	(128.85,162.91) --
	(129.59,162.62) --
	(130.32,162.62) --
	(131.06,162.77) --
	(131.80,161.46) --
	(132.54,161.46) --
	(133.27,162.19) --
	(134.01,161.75) --
	(134.75,161.03) --
	(135.49,162.19) --
	(136.23,161.75) --
	(136.96,160.59) --
	(137.70,160.88) --
	(138.44,160.30) --
	(139.18,160.30) --
	(139.91,160.45) --
	(140.65,160.45) --
	(141.39,160.30) --
	(142.13,160.45) --
	(142.87,160.74) --
	(143.60,161.03) --
	(144.34,161.90) --
	(145.08,162.48) --
	(145.82,161.75) --
	(146.55,161.17) --
	(147.29,160.59) --
	(148.03,160.01) --
	(148.77,161.32) --
	(149.51,162.19) --
	(150.24,162.48) --
	(150.98,162.19) --
	(151.72,162.48) --
	(152.46,162.91) --
	(153.19,161.61) --
	(153.93,161.90) --
	(154.67,160.59) --
	(155.41,161.61) --
	(156.15,160.59) --
	(156.88,161.32) --
	(157.62,161.17) --
	(158.36,162.62) --
	(159.10,161.61) --
	(159.83,161.90) --
	(160.57,161.46) --
	(161.31,161.61) --
	(162.05,160.59) --
	(162.78,160.88) --
	(163.52,161.03) --
	(164.26,161.17) --
	(165.00,161.46) --
	(165.74,161.46) --
	(166.47,160.45) --
	(167.21,160.16) --
	(167.95,159.87) --
	(168.69,160.74) --
	(169.42,160.88) --
	(170.16,160.74) --
	(170.90,161.03) --
	(171.64,161.61) --
	(172.38,160.88) --
	(173.11,161.90) --
	(173.85,162.77) --
	(174.59,162.33) --
	(175.33,162.04) --
	(176.06,161.90) --
	(176.80,161.90);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,151.92) rectangle (184.81,245.88);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,245.88) rectangle (184.81,253.83);

\path[] ( 54.44,245.88) rectangle (184.81,253.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 93.23,247.58) {T=250, K=3, c=0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,245.88) rectangle (184.81,253.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,159.72) -- (320.86,159.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,174.23) -- (320.86,174.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,188.74) -- (320.86,188.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,203.25) -- (320.86,203.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,217.76) -- (320.86,217.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,232.27) -- (320.86,232.27);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,156.42) {0.10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,170.93) {0.15};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,185.44) {0.20};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,199.95) {0.25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,214.46) {0.30};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,228.97) {0.35};
\end{scope}
\begin{scope}
\path[clip] (184.81,151.92) rectangle (315.17,245.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12,159.72) --
	(312.08,159.72);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (196.50,171.48) --
	(197.24,172.93) --
	(197.98,174.52) --
	(198.71,174.09) --
	(199.45,175.97) --
	(200.19,176.55) --
	(200.93,177.72) --
	(201.67,178.30) --
	(202.40,178.30) --
	(203.14,180.62) --
	(203.88,181.92) --
	(204.62,182.79) --
	(205.35,186.13) --
	(206.09,188.74) --
	(206.83,189.03) --
	(207.57,189.32) --
	(208.31,188.16) --
	(209.04,188.02) --
	(209.78,189.32) --
	(210.52,189.90) --
	(211.26,190.34) --
	(211.99,191.50) --
	(212.73,191.64) --
	(213.47,193.10) --
	(214.21,193.53) --
	(214.95,191.21) --
	(215.68,193.24) --
	(216.42,193.10) --
	(217.16,192.52) --
	(217.90,192.37) --
	(218.63,192.66) --
	(219.37,193.10) --
	(220.11,192.66) --
	(220.85,194.11) --
	(221.58,194.98) --
	(222.32,195.13) --
	(223.06,194.84) --
	(223.80,194.98) --
	(224.54,195.85) --
	(225.27,196.00) --
	(226.01,195.71) --
	(226.75,196.29) --
	(227.49,195.85) --
	(228.22,196.58) --
	(228.96,197.59) --
	(229.70,197.45) --
	(230.44,197.01) --
	(231.18,197.74) --
	(231.91,197.74) --
	(232.65,198.90) --
	(233.39,198.32) --
	(234.13,197.30) --
	(234.86,196.72) --
	(235.60,198.90) --
	(236.34,199.34) --
	(237.08,199.63) --
	(237.82,200.93) --
	(238.55,201.95) --
	(239.29,201.51) --
	(240.03,201.80) --
	(240.77,202.82) --
	(241.50,204.41) --
	(242.24,203.25) --
	(242.98,204.56) --
	(243.72,206.74) --
	(244.46,205.86) --
	(245.19,205.86) --
	(245.93,205.86) --
	(246.67,205.72) --
	(247.41,206.45) --
	(248.14,208.04) --
	(248.88,207.03) --
	(249.62,207.46) --
	(250.36,207.03) --
	(251.10,208.19) --
	(251.83,208.04) --
	(252.57,208.48) --
	(253.31,209.20) --
	(254.05,209.64) --
	(254.78,208.48) --
	(255.52,207.90) --
	(256.26,209.64) --
	(257.00,209.20) --
	(257.74,210.80) --
	(258.47,210.36) --
	(259.21,211.09) --
	(259.95,213.27) --
	(260.69,214.57) --
	(261.42,215.88) --
	(262.16,216.31) --
	(262.90,215.59) --
	(263.64,216.60) --
	(264.38,217.18) --
	(265.11,217.76) --
	(265.85,217.33) --
	(266.59,218.49) --
	(267.33,219.36) --
	(268.06,219.50) --
	(268.80,221.97) --
	(269.54,222.41) --
	(270.28,222.12) --
	(271.02,223.28) --
	(271.75,224.00) --
	(272.49,222.99) --
	(273.23,222.41) --
	(273.97,222.99) --
	(274.70,222.99) --
	(275.44,222.70) --
	(276.18,221.83) --
	(276.92,222.26) --
	(277.66,222.70) --
	(278.39,224.15) --
	(279.13,224.87) --
	(279.87,225.45) --
	(280.61,226.03) --
	(281.34,226.32) --
	(282.08,229.08) --
	(282.82,227.19) --
	(283.56,228.79) --
	(284.30,229.66) --
	(285.03,229.37) --
	(285.77,228.50) --
	(286.51,229.23) --
	(287.25,228.79) --
	(287.98,229.81) --
	(288.72,229.81) --
	(289.46,229.08) --
	(290.20,230.97) --
	(290.94,230.97) --
	(291.67,231.84) --
	(292.41,233.58) --
	(293.15,233.43) --
	(293.89,233.14) --
	(294.62,233.29) --
	(295.36,233.87) --
	(296.10,234.45) --
	(296.84,233.72) --
	(297.58,235.18) --
	(298.31,235.47) --
	(299.05,235.76) --
	(299.79,235.61) --
	(300.53,236.63) --
	(301.26,237.35) --
	(302.00,238.66) --
	(302.74,237.93) --
	(303.48,238.22) --
	(304.21,239.09) --
	(304.95,239.67) --
	(305.69,239.82) --
	(306.43,239.67) --
	(307.17,240.11);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,151.92) rectangle (315.17,245.88);
\end{scope}
\begin{scope}
\path[clip] (184.81,245.88) rectangle (315.17,253.83);

\path[] (184.81,245.88) rectangle (315.17,253.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (223.60,247.58) {T=250, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,245.88) rectangle (315.17,253.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76,355.73) -- ( 58.76,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65,355.73) -- ( 95.65,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (132.54,355.73) -- (132.54,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.42,355.73) -- (169.42,361.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,261.63) -- ( 48.75,261.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,276.14) -- ( 48.75,276.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,290.65) -- ( 48.75,290.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,305.16) -- ( 48.75,305.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,319.67) -- ( 48.75,319.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,334.18) -- ( 48.75,334.18);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,258.32) {0.10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,272.83) {0.15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,287.34) {0.20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,301.85) {0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,316.36) {0.30};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,330.87) {0.35};
\end{scope}
\begin{scope}
\path[clip] ( 54.44,253.83) rectangle (184.81,347.78);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76,261.63) --
	(181.72,261.63);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.20,270.33) --
	( 79.05,271.21) --
	( 80.89,271.35) --
	( 82.74,270.04) --
	( 84.58,271.06) --
	( 86.43,270.04) --
	( 88.27,269.17) --
	( 90.11,268.30) --
	( 91.96,269.03) --
	( 93.80,266.56) --
	( 95.65,268.16) --
	( 97.49,267.00) --
	( 99.34,266.42) --
	(101.18,266.71) --
	(103.03,264.82) --
	(104.87,264.97) --
	(106.71,265.98) --
	(108.56,265.84) --
	(110.40,266.42) --
	(112.25,265.98) --
	(114.09,266.27) --
	(115.94,265.40) --
	(117.78,265.26) --
	(119.63,265.40) --
	(121.47,265.11) --
	(123.31,265.98) --
	(125.16,266.27) --
	(127.00,265.40) --
	(128.85,265.40) --
	(130.69,266.85) --
	(132.54,265.69) --
	(134.38,266.13) --
	(136.23,265.55) --
	(138.07,263.95) --
	(139.91,264.97) --
	(141.76,263.95) --
	(143.60,264.10) --
	(145.45,263.66) --
	(147.29,264.82) --
	(149.14,265.69) --
	(150.98,265.40) --
	(152.83,264.97) --
	(154.67,263.81) --
	(156.51,263.66) --
	(158.36,262.93) --
	(160.20,263.22) --
	(162.05,263.51) --
	(163.89,263.66) --
	(165.74,264.10) --
	(167.58,263.22) --
	(169.42,263.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,253.83) rectangle (184.81,347.78);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,347.78) rectangle (184.81,355.73);

\path[] ( 54.44,347.78) rectangle (184.81,355.73);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 93.23,349.49) {T=100, K=3, c=0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,347.78) rectangle (184.81,355.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12,355.73) -- (189.12,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.01,355.73) -- (226.01,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.90,355.73) -- (262.90,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.79,355.73) -- (299.79,361.42);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (189.12,367.12) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (226.01,367.12) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (262.90,367.12) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (299.79,367.12) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,261.63) -- (320.86,261.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,276.14) -- (320.86,276.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,290.65) -- (320.86,290.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,305.16) -- (320.86,305.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,319.67) -- (320.86,319.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,334.18) -- (320.86,334.18);
\end{scope}
\begin{scope}
\path[clip] (184.81,253.83) rectangle (315.17,347.78);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12,261.63) --
	(312.08,261.63);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.57,289.20) --
	(209.41,287.31) --
	(211.26,288.76) --
	(213.10,287.46) --
	(214.95,288.91) --
	(216.79,292.10) --
	(218.63,293.26) --
	(220.48,293.99) --
	(222.32,294.42) --
	(224.17,295.73) --
	(226.01,296.31) --
	(227.86,296.89) --
	(229.70,299.21) --
	(231.54,299.94) --
	(233.39,300.52) --
	(235.23,301.39) --
	(237.08,302.26) --
	(238.92,302.84) --
	(240.77,305.16) --
	(242.61,306.47) --
	(244.46,305.74) --
	(246.30,306.03) --
	(248.14,306.32) --
	(249.99,307.63) --
	(251.83,308.21) --
	(253.68,307.19) --
	(255.52,309.37) --
	(257.37,309.80) --
	(259.21,310.38) --
	(261.06,312.85) --
	(262.90,312.99) --
	(264.74,315.61) --
	(266.59,316.62) --
	(268.43,318.36) --
	(270.28,318.51) --
	(272.12,320.25) --
	(273.97,319.81) --
	(275.81,320.83) --
	(277.66,320.98) --
	(279.50,322.57) --
	(281.34,324.60) --
	(283.19,324.31) --
	(285.03,324.75) --
	(286.88,326.92) --
	(288.72,325.76) --
	(290.57,326.92) --
	(292.41,326.49) --
	(294.26,327.80) --
	(296.10,327.80) --
	(297.94,329.10) --
	(299.79,332.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,253.83) rectangle (315.17,347.78);
\end{scope}
\begin{scope}
\path[clip] (184.81,347.78) rectangle (315.17,355.73);

\path[] (184.81,347.78) rectangle (315.17,355.73);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (223.60,349.49) {T=100, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,347.78) rectangle (315.17,355.73);
\end{scope}
\end{tikzpicture}