### 2017-04-21 floats/mc-mccrackensize.tex

% Created by tikzDevice version 0.6.2-92-0ad2792 on 2014-03-27 13:37:45
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor[named]{fillColor}{rgb}{1.00,1.00,1.00}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (361.35,397.48);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);

\path[] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at (184.81, 12.04) {$P/T$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.20] at ( 14.29,198.90) {Rejection Probability};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 59.11) -- ( 48.75, 59.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 79.93) -- ( 48.75, 79.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,100.74) -- ( 48.75,100.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,121.55) -- ( 48.75,121.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,142.37) -- ( 48.75,142.37);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 55.81) {0.10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 76.62) {0.15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06, 97.43) {0.20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,118.25) {0.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76, 50.02) -- ( 58.76, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65, 50.02) -- ( 95.65, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (132.54, 50.02) -- (132.54, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.42, 50.02) -- (169.42, 44.32);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 58.76, 32.02) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 95.65, 32.02) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (132.54, 32.02) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (169.42, 32.02) {0.6};
\end{scope}
\begin{scope}
\path[clip] ( 54.44, 50.02) rectangle (184.81,143.97);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76, 59.11) --
	(181.72, 59.11);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.45, 61.40) --
	( 62.82, 62.65) --
	( 63.19, 63.90) --
	( 63.56, 61.40) --
	( 63.92, 61.61) --
	( 64.29, 64.73) --
	( 64.66, 66.19) --
	( 65.03, 63.49) --
	( 65.40, 64.53) --
	( 65.77, 62.24) --
	( 66.14, 63.69) --
	( 66.51, 60.78) --
	( 66.88, 64.94) --
	( 67.24, 62.24) --
	( 67.61, 63.69) --
	( 67.98, 66.19) --
	( 68.35, 62.65) --
	( 68.72, 61.61) --
	( 69.09, 63.49) --
	( 69.46, 66.40) --
	( 69.83, 63.28) --
	( 70.20, 63.90) --
	( 70.56, 66.19) --
	( 70.93, 64.94) --
	( 71.30, 66.82) --
	( 71.67, 68.06) --
	( 72.04, 66.61) --
	( 72.41, 68.48) --
	( 72.78, 67.44) --
	( 73.15, 66.40) --
	( 73.52, 66.40) --
	( 73.88, 66.19) --
	( 74.25, 64.94) --
	( 74.62, 66.19) --
	( 74.99, 65.57) --
	( 75.36, 65.15) --
	( 75.73, 64.53) --
	( 76.10, 62.03) --
	( 76.47, 63.28) --
	( 76.84, 64.11) --
	( 77.20, 63.28) --
	( 77.57, 60.78) --
	( 77.94, 59.95) --
	( 78.31, 61.61) --
	( 78.68, 63.49) --
	( 79.05, 62.86) --
	( 79.42, 62.44) --
	( 79.79, 61.40) --
	( 80.16, 61.82) --
	( 80.52, 62.86) --
	( 80.89, 63.49) --
	( 81.26, 61.40) --
	( 81.63, 62.86) --
	( 82.00, 63.69) --
	( 82.37, 62.65) --
	( 82.74, 63.49) --
	( 83.11, 65.98) --
	( 83.48, 65.15) --
	( 83.84, 63.28) --
	( 84.21, 63.28) --
	( 84.58, 65.36) --
	( 84.95, 65.15) --
	( 85.32, 66.19) --
	( 85.69, 66.19) --
	( 86.06, 66.19) --
	( 86.43, 65.77) --
	( 86.79, 67.86) --
	( 87.16, 66.40) --
	( 87.53, 67.65) --
	( 87.90, 69.10) --
	( 88.27, 68.90) --
	( 88.64, 69.10) --
	( 89.01, 67.44) --
	( 89.38, 67.02) --
	( 89.75, 68.48) --
	( 90.11, 67.44) --
	( 90.48, 67.44) --
	( 90.85, 68.90) --
	( 91.22, 68.90) --
	( 91.59, 68.27) --
	( 91.96, 68.69) --
	( 92.33, 69.73) --
	( 92.70, 69.52) --
	( 93.07, 69.31) --
	( 93.43, 67.65) --
	( 93.80, 67.44) --
	( 94.17, 68.06) --
	( 94.54, 67.23) --
	( 94.91, 66.19) --
	( 95.28, 65.15) --
	( 95.65, 65.15) --
	( 96.02, 66.19) --
	( 96.39, 65.36) --
	( 96.75, 67.65) --
	( 97.12, 66.40) --
	( 97.49, 65.57) --
	( 97.86, 65.15) --
	( 98.23, 65.15) --
	( 98.60, 63.69) --
	( 98.97, 62.44) --
	( 99.34, 63.69) --
	( 99.71, 63.69) --
	(100.07, 63.28) --
	(100.44, 62.24) --
	(100.81, 61.82) --
	(101.18, 59.95) --
	(101.55, 61.82) --
	(101.92, 62.44) --
	(102.29, 60.57) --
	(102.66, 61.40) --
	(103.03, 61.82) --
	(103.39, 62.24) --
	(103.76, 62.44) --
	(104.13, 62.03) --
	(104.50, 60.78) --
	(104.87, 59.11) --
	(105.24, 60.99) --
	(105.61, 62.03) --
	(105.98, 63.49) --
	(106.35, 61.61) --
	(106.71, 61.82) --
	(107.08, 59.74) --
	(107.45, 59.53) --
	(107.82, 60.57) --
	(108.19, 58.91) --
	(108.56, 61.40) --
	(108.93, 60.16) --
	(109.30, 60.16) --
	(109.67, 60.57) --
	(110.03, 61.82) --
	(110.40, 61.20) --
	(110.77, 61.40) --
	(111.14, 61.20) --
	(111.51, 61.20) --
	(111.88, 61.61) --
	(112.25, 61.61) --
	(112.62, 61.40) --
	(112.99, 60.57) --
	(113.35, 62.03) --
	(113.72, 61.61) --
	(114.09, 62.24) --
	(114.46, 62.24) --
	(114.83, 62.03) --
	(115.20, 64.11) --
	(115.57, 62.24) --
	(115.94, 59.74) --
	(116.31, 59.53) --
	(116.67, 58.28) --
	(117.04, 57.87) --
	(117.41, 58.49) --
	(117.78, 59.53) --
	(118.15, 59.95) --
	(118.52, 58.28) --
	(118.89, 59.32) --
	(119.26, 59.95) --
	(119.63, 59.53) --
	(119.99, 59.53) --
	(120.36, 59.53) --
	(120.73, 59.95) --
	(121.10, 59.53) --
	(121.47, 59.74) --
	(121.84, 59.74) --
	(122.21, 58.70) --
	(122.58, 56.41) --
	(122.95, 61.20) --
	(123.31, 60.57) --
	(123.68, 58.07) --
	(124.05, 59.11) --
	(124.42, 58.91) --
	(124.79, 58.49) --
	(125.16, 58.91) --
	(125.53, 59.32) --
	(125.90, 59.32) --
	(126.27, 59.53) --
	(126.63, 58.70) --
	(127.00, 60.78) --
	(127.37, 59.74) --
	(127.74, 58.49) --
	(128.11, 60.16) --
	(128.48, 58.91) --
	(128.85, 58.91) --
	(129.22, 59.32) --
	(129.59, 60.36) --
	(129.95, 60.78) --
	(130.32, 59.53) --
	(130.69, 59.11) --
	(131.06, 62.03) --
	(131.43, 62.86) --
	(131.80, 62.65) --
	(132.17, 63.49) --
	(132.54, 63.69) --
	(132.91, 62.86) --
	(133.27, 61.20) --
	(133.64, 60.16) --
	(134.01, 62.86) --
	(134.38, 62.65) --
	(134.75, 62.86) --
	(135.12, 62.86) --
	(135.49, 60.99) --
	(135.86, 59.74) --
	(136.23, 60.57) --
	(136.59, 61.82) --
	(136.96, 62.86) --
	(137.33, 63.28) --
	(137.70, 63.28) --
	(138.07, 63.69) --
	(138.44, 61.40) --
	(138.81, 60.57) --
	(139.18, 60.16) --
	(139.55, 59.95) --
	(139.91, 59.53) --
	(140.28, 60.16) --
	(140.65, 62.03) --
	(141.02, 60.78) --
	(141.39, 61.82) --
	(141.76, 61.20) --
	(142.13, 62.44) --
	(142.50, 61.40) --
	(142.87, 60.57) --
	(143.23, 59.53) --
	(143.60, 60.16) --
	(143.97, 59.74) --
	(144.34, 60.57) --
	(144.71, 60.78) --
	(145.08, 62.03) --
	(145.45, 60.36) --
	(145.82, 60.16) --
	(146.19, 61.20) --
	(146.55, 61.40) --
	(146.92, 59.95) --
	(147.29, 60.36) --
	(147.66, 59.53) --
	(148.03, 59.74) --
	(148.40, 56.62) --
	(148.77, 55.78) --
	(149.14, 56.41) --
	(149.51, 59.53) --
	(149.87, 58.28) --
	(150.24, 58.49) --
	(150.61, 57.87) --
	(150.98, 59.11) --
	(151.35, 58.49) --
	(151.72, 56.83) --
	(152.09, 56.41) --
	(152.46, 59.32) --
	(152.83, 59.32) --
	(153.19, 59.74) --
	(153.56, 57.87) --
	(153.93, 59.11) --
	(154.30, 58.70) --
	(154.67, 57.87) --
	(155.04, 58.07) --
	(155.41, 56.83) --
	(155.78, 57.03) --
	(156.15, 58.70) --
	(156.51, 60.36) --
	(156.88, 59.11) --
	(157.25, 62.03) --
	(157.62, 61.40) --
	(157.99, 60.57) --
	(158.36, 62.24) --
	(158.73, 61.61) --
	(159.10, 59.74) --
	(159.47, 61.61) --
	(159.83, 60.36) --
	(160.20, 61.61) --
	(160.57, 59.32) --
	(160.94, 61.20) --
	(161.31, 59.11) --
	(161.68, 58.28) --
	(162.05, 60.16) --
	(162.42, 60.16) --
	(162.78, 59.11) --
	(163.15, 57.66) --
	(163.52, 59.53) --
	(163.89, 58.91) --
	(164.26, 59.53) --
	(164.63, 59.11) --
	(165.00, 58.49) --
	(165.37, 59.32) --
	(165.74, 59.32) --
	(166.10, 58.91) --
	(166.47, 59.11) --
	(166.84, 57.66) --
	(167.21, 59.74) --
	(167.58, 59.95) --
	(167.95, 58.70) --
	(168.32, 59.32) --
	(168.69, 58.91) --
	(169.06, 58.49) --
	(169.42, 55.78) --
	(169.79, 56.83) --
	(170.16, 58.91) --
	(170.53, 58.07) --
	(170.90, 59.11) --
	(171.27, 57.66) --
	(171.64, 57.24) --
	(172.01, 57.66) --
	(172.38, 57.24) --
	(172.74, 55.99) --
	(173.11, 58.28) --
	(173.48, 55.99) --
	(173.85, 57.87) --
	(174.22, 59.95) --
	(174.59, 58.49) --
	(174.96, 59.32) --
	(175.33, 58.49) --
	(175.70, 59.74) --
	(176.06, 59.53) --
	(176.43, 62.65) --
	(176.80, 63.07);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44, 50.02) rectangle (184.81,143.97);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,143.97) rectangle (184.81,151.92);

\path[] ( 54.44,143.97) rectangle (184.81,151.92);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 93.23,145.67) {T=500, K=3, c=0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,143.97) rectangle (184.81,151.92);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12, 50.02) -- (189.12, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.01, 50.02) -- (226.01, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.90, 50.02) -- (262.90, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.79, 50.02) -- (299.79, 44.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17, 59.11) -- (320.86, 59.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17, 79.93) -- (320.86, 79.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,100.74) -- (320.86,100.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,121.55) -- (320.86,121.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,142.37) -- (320.86,142.37);
\end{scope}
\begin{scope}
\path[clip] (184.81, 50.02) rectangle (315.17,143.97);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12, 59.11) --
	(312.08, 59.11);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (192.81, 66.19) --
	(193.18, 66.61) --
	(193.55, 68.48) --
	(193.92, 66.82) --
	(194.29, 65.77) --
	(194.66, 67.44) --
	(195.03, 66.19) --
	(195.39, 69.73) --
	(195.76, 68.27) --
	(196.13, 66.61) --
	(196.50, 64.94) --
	(196.87, 67.02) --
	(197.24, 69.31) --
	(197.61, 68.69) --
	(197.98, 67.86) --
	(198.35, 69.52) --
	(198.71, 69.10) --
	(199.08, 67.65) --
	(199.45, 68.27) --
	(199.82, 67.44) --
	(200.19, 68.90) --
	(200.56, 70.98) --
	(200.93, 69.52) --
	(201.30, 73.48) --
	(201.67, 73.27) --
	(202.03, 72.85) --
	(202.40, 73.89) --
	(202.77, 75.35) --
	(203.14, 74.52) --
	(203.51, 74.93) --
	(203.88, 75.97) --
	(204.25, 76.60) --
	(204.62, 75.76) --
	(204.99, 77.64) --
	(205.35, 75.76) --
	(205.72, 77.01) --
	(206.09, 80.55) --
	(206.46, 80.55) --
	(206.83, 79.72) --
	(207.20, 81.80) --
	(207.57, 81.18) --
	(207.94, 77.43) --
	(208.31, 79.93) --
	(208.67, 80.34) --
	(209.04, 81.59) --
	(209.41, 80.97) --
	(209.78, 81.38) --
	(210.15, 80.55) --
	(210.52, 80.55) --
	(210.89, 81.80) --
	(211.26, 84.09) --
	(211.63, 83.67) --
	(211.99, 85.13) --
	(212.36, 85.13) --
	(212.73, 84.92) --
	(213.10, 84.09) --
	(213.47, 85.75) --
	(213.84, 84.51) --
	(214.21, 84.71) --
	(214.58, 86.80) --
	(214.95, 86.38) --
	(215.31, 87.84) --
	(215.68, 89.29) --
	(216.05, 87.00) --
	(216.42, 88.04) --
	(216.79, 89.29) --
	(217.16, 87.42) --
	(217.53, 87.21) --
	(217.90, 87.21) --
	(218.27, 89.50) --
	(218.63, 88.04) --
	(219.00, 85.96) --
	(219.37, 86.80) --
	(219.74, 86.80) --
	(220.11, 87.00) --
	(220.48, 85.75) --
	(220.85, 90.13) --
	(221.22, 92.83) --
	(221.58, 92.83) --
	(221.95, 92.62) --
	(222.32, 91.37) --
	(222.69, 90.33) --
	(223.06, 90.54) --
	(223.43, 92.41) --
	(223.80, 90.54) --
	(224.17, 88.88) --
	(224.54, 88.46) --
	(224.90, 87.42) --
	(225.27, 90.13) --
	(225.64, 90.13) --
	(226.01, 90.96) --
	(226.38, 92.41) --
	(226.75, 93.66) --
	(227.12, 90.13) --
	(227.49, 91.37) --
	(227.86, 94.29) --
	(228.22, 95.12) --
	(228.59, 93.04) --
	(228.96, 93.66) --
	(229.33, 93.87) --
	(229.70, 95.75) --
	(230.07, 94.29) --
	(230.44, 93.87) --
	(230.81, 94.29) --
	(231.18, 95.33) --
	(231.54, 93.25) --
	(231.91, 95.95) --
	(232.28, 94.91) --
	(232.65, 95.75) --
	(233.02, 93.66) --
	(233.39, 93.04) --
	(233.76, 94.29) --
	(234.13, 95.54) --
	(234.50, 97.20) --
	(234.86, 97.41) --
	(235.23, 97.41) --
	(235.60, 98.24) --
	(235.97, 97.41) --
	(236.34, 97.83) --
	(236.71, 98.24) --
	(237.08, 97.20) --
	(237.45, 98.45) --
	(237.82, 99.91) --
	(238.18, 98.45) --
	(238.55, 97.83) --
	(238.92, 95.95) --
	(239.29, 96.99) --
	(239.66, 97.41) --
	(240.03, 95.95) --
	(240.40, 97.41) --
	(240.77, 98.66) --
	(241.14, 98.66) --
	(241.50, 98.66) --
	(241.87, 99.08) --
	(242.24, 98.24) --
	(242.61, 97.83) --
	(242.98, 99.49) --
	(243.35, 97.83) --
	(243.72, 97.62) --
	(244.09, 96.79) --
	(244.46, 96.58) --
	(244.82, 97.41) --
	(245.19, 98.66) --
	(245.56, 98.24) --
	(245.93, 99.49) --
	(246.30, 99.49) --
	(246.67, 99.49) --
	(247.04, 99.08) --
	(247.41, 99.70) --
	(247.78, 99.91) --
	(248.14,101.78) --
	(248.51,101.78) --
	(248.88,104.49) --
	(249.25,102.82) --
	(249.62,103.65) --
	(249.99,103.65) --
	(250.36,103.45) --
	(250.73,105.53) --
	(251.10,102.61) --
	(251.46,105.53) --
	(251.83,105.11) --
	(252.20,104.07) --
	(252.57,105.53) --
	(252.94,103.65) --
	(253.31,103.24) --
	(253.68,103.65) --
	(254.05,103.65) --
	(254.42,103.45) --
	(254.78,105.32) --
	(255.15,106.78) --
	(255.52,105.94) --
	(255.89,106.15) --
	(256.26,106.78) --
	(256.63,107.19) --
	(257.00,107.61) --
	(257.37,106.78) --
	(257.74,107.40) --
	(258.10,107.40) --
	(258.47,108.44) --
	(258.84,109.07) --
	(259.21,111.56) --
	(259.58,110.94) --
	(259.95,109.07) --
	(260.32,111.98) --
	(260.69,110.52) --
	(261.06,109.69) --
	(261.42,110.73) --
	(261.79,110.73) --
	(262.16,112.19) --
	(262.53,111.98) --
	(262.90,110.94) --
	(263.27,110.73) --
	(263.64,110.94) --
	(264.01,111.77) --
	(264.38,112.81) --
	(264.74,112.81) --
	(265.11,111.98) --
	(265.48,111.77) --
	(265.85,111.56) --
	(266.22,111.56) --
	(266.59,112.81) --
	(266.96,113.85) --
	(267.33,114.27) --
	(267.70,114.06) --
	(268.06,115.73) --
	(268.43,115.73) --
	(268.80,117.60) --
	(269.17,119.26) --
	(269.54,116.35) --
	(269.91,116.35) --
	(270.28,116.97) --
	(270.65,117.39) --
	(271.02,115.31) --
	(271.38,114.89) --
	(271.75,114.06) --
	(272.12,114.68) --
	(272.49,115.31) --
	(272.86,116.77) --
	(273.23,113.23) --
	(273.60,115.73) --
	(273.97,116.77) --
	(274.34,115.73) --
	(274.70,116.35) --
	(275.07,118.01) --
	(275.44,119.68) --
	(275.81,120.10) --
	(276.18,119.26) --
	(276.55,119.89) --
	(276.92,120.93) --
	(277.29,122.18) --
	(277.66,122.39) --
	(278.02,122.59) --
	(278.39,122.39) --
	(278.76,124.05) --
	(279.13,122.39) --
	(279.50,122.18) --
	(279.87,122.18) --
	(280.24,123.01) --
	(280.61,123.63) --
	(280.98,125.09) --
	(281.34,125.92) --
	(281.71,127.59) --
	(282.08,125.51) --
	(282.45,125.09) --
	(282.82,124.88) --
	(283.19,127.38) --
	(283.56,125.51) --
	(283.93,126.96) --
	(284.30,126.34) --
	(284.66,125.72) --
	(285.03,124.88) --
	(285.40,125.30) --
	(285.77,125.51) --
	(286.14,125.09) --
	(286.51,122.80) --
	(286.88,122.39) --
	(287.25,122.59) --
	(287.62,124.67) --
	(287.98,124.67) --
	(288.35,125.72) --
	(288.72,126.55) --
	(289.09,126.96) --
	(289.46,126.76) --
	(289.83,125.09) --
	(290.20,124.47) --
	(290.57,126.55) --
	(290.94,127.59) --
	(291.30,126.76) --
	(291.67,126.76) --
	(292.04,126.96) --
	(292.41,128.00) --
	(292.78,129.46) --
	(293.15,128.63) --
	(293.52,127.17) --
	(293.89,128.21) --
	(294.26,129.05) --
	(294.62,128.42) --
	(294.99,127.59) --
	(295.36,127.17) --
	(295.73,128.42) --
	(296.10,129.88) --
	(296.47,128.21) --
	(296.84,126.96) --
	(297.21,127.17) --
	(297.58,124.88) --
	(297.94,124.47) --
	(298.31,124.05) --
	(298.68,124.47) --
	(299.05,127.17) --
	(299.42,124.88) --
	(299.79,125.72) --
	(300.16,126.13) --
	(300.53,124.47) --
	(300.89,127.38) --
	(301.26,126.13) --
	(301.63,124.67) --
	(302.00,125.51) --
	(302.37,124.05) --
	(302.74,122.80) --
	(303.11,125.72) --
	(303.48,127.38) --
	(303.85,128.21) --
	(304.21,129.25) --
	(304.58,125.72) --
	(304.95,125.72) --
	(305.32,125.92) --
	(305.69,127.17) --
	(306.06,127.38) --
	(306.43,128.42) --
	(306.80,129.46) --
	(307.17,128.63);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81, 50.02) rectangle (315.17,143.97);
\end{scope}
\begin{scope}
\path[clip] (184.81,143.97) rectangle (315.17,151.92);

\path[] (184.81,143.97) rectangle (315.17,151.92);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (223.60,145.67) {T=500, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,143.97) rectangle (315.17,151.92);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,161.02) -- ( 48.75,161.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,181.83) -- ( 48.75,181.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,202.65) -- ( 48.75,202.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,223.46) -- ( 48.75,223.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,244.27) -- ( 48.75,244.27);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,151.92) rectangle (184.81,245.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76,161.02) --
	(181.72,161.02);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.14,164.56) --
	( 66.88,165.39) --
	( 67.61,163.73) --
	( 68.35,164.14) --
	( 69.09,163.93) --
	( 69.83,168.10) --
	( 70.56,168.93) --
	( 71.30,168.31) --
	( 72.04,173.09) --
	( 72.78,171.01) --
	( 73.52,169.55) --
	( 74.25,168.72) --
	( 74.99,167.26) --
	( 75.73,169.35) --
	( 76.47,170.80) --
	( 77.20,171.22) --
	( 77.94,171.01) --
	( 78.68,169.76) --
	( 79.42,166.43) --
	( 80.16,167.47) --
	( 80.89,167.89) --
	( 81.63,165.39) --
	( 82.37,167.06) --
	( 83.11,167.47) --
	( 83.84,166.64) --
	( 84.58,166.43) --
	( 85.32,166.85) --
	( 86.06,168.51) --
	( 86.79,169.14) --
	( 87.53,169.76) --
	( 88.27,165.60) --
	( 89.01,167.47) --
	( 89.75,170.39) --
	( 90.48,168.31) --
	( 91.22,169.97) --
	( 91.96,170.39) --
	( 92.70,167.26) --
	( 93.43,167.47) --
	( 94.17,165.81) --
	( 94.91,165.39) --
	( 95.65,166.22) --
	( 96.39,166.64) --
	( 97.12,168.72) --
	( 97.86,166.85) --
	( 98.60,166.64) --
	( 99.34,166.85) --
	(100.07,167.06) --
	(100.81,168.72) --
	(101.55,167.47) --
	(102.29,167.26) --
	(103.03,165.60) --
	(103.76,164.35) --
	(104.50,163.93) --
	(105.24,164.35) --
	(105.98,165.60) --
	(106.71,166.22) --
	(107.45,165.18) --
	(108.19,162.89) --
	(108.93,164.35) --
	(109.67,163.52) --
	(110.40,163.31) --
	(111.14,162.06) --
	(111.88,163.10) --
	(112.62,164.14) --
	(113.35,163.10) --
	(114.09,162.27) --
	(114.83,163.52) --
	(115.57,162.89) --
	(116.31,163.31) --
	(117.04,162.69) --
	(117.78,164.35) --
	(118.52,163.73) --
	(119.26,162.89) --
	(119.99,162.06) --
	(120.73,162.89) --
	(121.47,163.10) --
	(122.21,162.89) --
	(122.95,161.65) --
	(123.68,164.14) --
	(124.42,164.56) --
	(125.16,162.69) --
	(125.90,163.10) --
	(126.63,162.27) --
	(127.37,161.85) --
	(128.11,162.27) --
	(128.85,162.69) --
	(129.59,162.48) --
	(130.32,163.73) --
	(131.06,162.69) --
	(131.80,160.60) --
	(132.54,163.52) --
	(133.27,161.44) --
	(134.01,159.77) --
	(134.75,161.44) --
	(135.49,162.06) --
	(136.23,162.48) --
	(136.96,162.69) --
	(137.70,160.40) --
	(138.44,160.81) --
	(139.18,164.14) --
	(139.91,164.56) --
	(140.65,164.14) --
	(141.39,162.69) --
	(142.13,160.40) --
	(142.87,162.06) --
	(143.60,161.44) --
	(144.34,161.23) --
	(145.08,162.27) --
	(145.82,162.48) --
	(146.55,161.85) --
	(147.29,162.48) --
	(148.03,162.48) --
	(148.77,163.73) --
	(149.51,164.35) --
	(150.24,166.43) --
	(150.98,167.68) --
	(151.72,164.77) --
	(152.46,166.22) --
	(153.19,165.18) --
	(153.93,166.85) --
	(154.67,168.72) --
	(155.41,166.85) --
	(156.15,165.81) --
	(156.88,163.73) --
	(157.62,163.31) --
	(158.36,162.89) --
	(159.10,162.27) --
	(159.83,162.06) --
	(160.57,164.14) --
	(161.31,164.35) --
	(162.05,163.10) --
	(162.78,163.93) --
	(163.52,164.56) --
	(164.26,161.65) --
	(165.00,162.89) --
	(165.74,166.43) --
	(166.47,162.69) --
	(167.21,164.35) --
	(167.95,163.52) --
	(168.69,161.65) --
	(169.42,160.81) --
	(170.16,162.27) --
	(170.90,161.44) --
	(171.64,161.44) --
	(172.38,160.81) --
	(173.11,162.06) --
	(173.85,161.85) --
	(174.59,162.06) --
	(175.33,165.60) --
	(176.06,163.73) --
	(176.80,165.60);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,151.92) rectangle (184.81,245.88);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,245.88) rectangle (184.81,253.83);

\path[] ( 54.44,245.88) rectangle (184.81,253.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 93.23,247.58) {T=250, K=3, c=0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,245.88) rectangle (184.81,253.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,161.02) -- (320.86,161.02);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,181.83) -- (320.86,181.83);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,202.65) -- (320.86,202.65);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,223.46) -- (320.86,223.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,244.27) -- (320.86,244.27);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,157.72) {0.10};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,178.53) {0.15};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,199.34) {0.20};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.96] at (326.55,220.15) {0.25};
\end{scope}
\begin{scope}
\path[clip] (184.81,151.92) rectangle (315.17,245.88);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12,161.02) --
	(312.08,161.02);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (196.50,170.80) --
	(197.24,171.64) --
	(197.98,172.88) --
	(198.71,172.26) --
	(199.45,174.76) --
	(200.19,176.21) --
	(200.93,177.67) --
	(201.67,179.34) --
	(202.40,180.58) --
	(203.14,179.75) --
	(203.88,178.09) --
	(204.62,180.58) --
	(205.35,182.46) --
	(206.09,185.79) --
	(206.83,187.24) --
	(207.57,189.12) --
	(208.31,189.53) --
	(209.04,187.66) --
	(209.78,189.12) --
	(210.52,189.74) --
	(211.26,189.33) --
	(211.99,192.24) --
	(212.73,191.20) --
	(213.47,194.11) --
	(214.21,195.57) --
	(214.95,195.57) --
	(215.68,196.40) --
	(216.42,194.95) --
	(217.16,193.28) --
	(217.90,190.37) --
	(218.63,193.07) --
	(219.37,191.20) --
	(220.11,193.07) --
	(220.85,193.90) --
	(221.58,194.11) --
	(222.32,197.03) --
	(223.06,197.44) --
	(223.80,196.40) --
	(224.54,198.07) --
	(225.27,197.23) --
	(226.01,197.03) --
	(226.75,197.03) --
	(227.49,197.86) --
	(228.22,198.48) --
	(228.96,198.48) --
	(229.70,199.94) --
	(230.44,199.94) --
	(231.18,204.94) --
	(231.91,204.52) --
	(232.65,203.06) --
	(233.39,201.40) --
	(234.13,201.19) --
	(234.86,197.65) --
	(235.60,198.69) --
	(236.34,199.11) --
	(237.08,199.52) --
	(237.82,199.73) --
	(238.55,199.94) --
	(239.29,200.98) --
	(240.03,200.77) --
	(240.77,202.65) --
	(241.50,204.10) --
	(242.24,206.18) --
	(242.98,207.85) --
	(243.72,210.35) --
	(244.46,209.31) --
	(245.19,209.31) --
	(245.93,209.93) --
	(246.67,208.89) --
	(247.41,210.14) --
	(248.14,210.76) --
	(248.88,211.80) --
	(249.62,210.56) --
	(250.36,210.76) --
	(251.10,212.01) --
	(251.83,211.39) --
	(252.57,209.72) --
	(253.31,210.56) --
	(254.05,211.18) --
	(254.78,212.01) --
	(255.52,212.84) --
	(256.26,213.89) --
	(257.00,213.47) --
	(257.74,214.51) --
	(258.47,214.93) --
	(259.21,218.46) --
	(259.95,218.05) --
	(260.69,221.79) --
	(261.42,218.88) --
	(262.16,219.30) --
	(262.90,219.71) --
	(263.64,221.17) --
	(264.38,223.04) --
	(265.11,222.00) --
	(265.85,224.08) --
	(266.59,224.92) --
	(267.33,226.79) --
	(268.06,227.83) --
	(268.80,225.54) --
	(269.54,227.62) --
	(270.28,229.08) --
	(271.02,228.45) --
	(271.75,228.87) --
	(272.49,229.49) --
	(273.23,230.33) --
	(273.97,229.49) --
	(274.70,231.37) --
	(275.44,229.70) --
	(276.18,229.91) --
	(276.92,229.91) --
	(277.66,230.33) --
	(278.39,230.33) --
	(279.13,230.74) --
	(279.87,230.54) --
	(280.61,231.37) --
	(281.34,231.16) --
	(282.08,232.20) --
	(282.82,233.66) --
	(283.56,232.62) --
	(284.30,234.28) --
	(285.03,234.70) --
	(285.77,234.28) --
	(286.51,233.03) --
	(287.25,233.03) --
	(287.98,233.03) --
	(288.72,235.11) --
	(289.46,233.87) --
	(290.20,232.62) --
	(290.94,234.07) --
	(291.67,234.70) --
	(292.41,234.28) --
	(293.15,232.62) --
	(293.89,235.53) --
	(294.62,234.91) --
	(295.36,235.95) --
	(296.10,234.91) --
	(296.84,236.57) --
	(297.58,235.11) --
	(298.31,235.53) --
	(299.05,233.03) --
	(299.79,235.11) --
	(300.53,234.28) --
	(301.26,233.45) --
	(302.00,232.62) --
	(302.74,230.95) --
	(303.48,233.03) --
	(304.21,234.49) --
	(304.95,235.95) --
	(305.69,238.44) --
	(306.43,240.11) --
	(307.17,238.86);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,151.92) rectangle (315.17,245.88);
\end{scope}
\begin{scope}
\path[clip] (184.81,245.88) rectangle (315.17,253.83);

\path[] (184.81,245.88) rectangle (315.17,253.83);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (223.60,247.58) {T=250, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,245.88) rectangle (315.17,253.83);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76,355.73) -- ( 58.76,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 95.65,355.73) -- ( 95.65,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (132.54,355.73) -- (132.54,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (169.42,355.73) -- (169.42,361.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,262.93) -- ( 48.75,262.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,283.74) -- ( 48.75,283.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,304.55) -- ( 48.75,304.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,325.37) -- ( 48.75,325.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,346.18) -- ( 48.75,346.18);

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,259.62) {0.10};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,280.43) {0.15};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,301.25) {0.20};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.96] at ( 43.06,322.06) {0.25};
\end{scope}
\begin{scope}
\path[clip] ( 54.44,253.83) rectangle (184.81,347.78);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 58.76,262.93) --
	(181.72,262.93);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 77.20,265.63) --
	( 79.05,265.01) --
	( 80.89,268.13) --
	( 82.74,265.84) --
	( 84.58,267.71) --
	( 86.43,267.09) --
	( 88.27,266.47) --
	( 90.11,268.34) --
	( 91.96,268.55) --
	( 93.80,265.01) --
	( 95.65,267.71) --
	( 97.49,267.30) --
	( 99.34,265.63) --
	(101.18,263.76) --
	(103.03,264.38) --
	(104.87,263.97) --
	(106.71,265.84) --
	(108.56,264.59) --
	(110.40,265.22) --
	(112.25,265.84) --
	(114.09,267.09) --
	(115.94,263.14) --
	(117.78,265.01) --
	(119.63,264.18) --
	(121.47,263.97) --
	(123.31,265.84) --
	(125.16,263.55) --
	(127.00,262.72) --
	(128.85,262.93) --
	(130.69,264.80) --
	(132.54,267.09) --
	(134.38,267.51) --
	(136.23,268.13) --
	(138.07,269.17) --
	(139.91,268.34) --
	(141.76,269.38) --
	(143.60,267.30) --
	(145.45,264.59) --
	(147.29,267.30) --
	(149.14,266.47) --
	(150.98,269.80) --
	(152.83,266.05) --
	(154.67,265.63) --
	(156.51,266.47) --
	(158.36,269.80) --
	(160.20,269.80) --
	(162.05,270.21) --
	(163.89,269.80) --
	(165.74,267.09) --
	(167.58,266.47) --
	(169.42,264.59);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,253.83) rectangle (184.81,347.78);
\end{scope}
\begin{scope}
\path[clip] ( 54.44,347.78) rectangle (184.81,355.73);

\path[] ( 54.44,347.78) rectangle (184.81,355.73);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at ( 93.23,349.49) {T=100, K=3, c=0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 54.44,347.78) rectangle (184.81,355.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12,355.73) -- (189.12,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (226.01,355.73) -- (226.01,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.90,355.73) -- (262.90,361.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.79,355.73) -- (299.79,361.42);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (189.12,367.12) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (226.01,367.12) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (262.90,367.12) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.96] at (299.79,367.12) {0.6};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,262.93) -- (320.86,262.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,283.74) -- (320.86,283.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,304.55) -- (320.86,304.55);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,325.37) -- (320.86,325.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (315.17,346.18) -- (320.86,346.18);
\end{scope}
\begin{scope}
\path[clip] (184.81,253.83) rectangle (315.17,347.78);
\definecolor[named]{drawColor}{rgb}{0.83,0.83,0.83}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.12,262.93) --
	(312.08,262.93);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (207.57,284.78) --
	(209.41,283.12) --
	(211.26,284.36) --
	(213.10,286.86) --
	(214.95,287.28) --
	(216.79,290.40) --
	(218.63,289.78) --
	(220.48,291.44) --
	(222.32,294.15) --
	(224.17,294.56) --
	(226.01,294.98) --
	(227.86,297.89) --
	(229.70,299.14) --
	(231.54,299.77) --
	(233.39,297.89) --
	(235.23,299.35) --
	(237.08,301.43) --
	(238.92,306.84) --
	(240.77,307.05) --
	(242.61,307.26) --
	(244.46,306.63) --
	(246.30,306.84) --
	(248.14,309.96) --
	(249.99,310.80) --
	(251.83,311.21) --
	(253.68,311.42) --
	(255.52,313.50) --
	(257.37,311.84) --
	(259.21,312.67) --
	(261.06,314.13) --
	(262.90,316.00) --
	(264.74,319.75) --
	(266.59,321.83) --
	(268.43,319.75) --
	(270.28,322.24) --
	(272.12,321.20) --
	(273.97,320.58) --
	(275.81,322.24) --
	(277.66,325.99) --
	(279.50,326.61) --
	(281.34,330.36) --
	(283.19,331.19) --
	(285.03,329.11) --
	(286.88,329.94) --
	(288.72,328.28) --
	(290.57,329.53) --
	(292.41,329.94) --
	(294.26,329.53) --
	(296.10,332.03) --
	(297.94,330.98) --
	(299.79,332.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,253.83) rectangle (315.17,347.78);
\end{scope}
\begin{scope}
\path[clip] (184.81,347.78) rectangle (315.17,355.73);

\path[] (184.81,347.78) rectangle (315.17,355.73);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.66] at (223.60,349.49) {T=100, K=3, c=1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,397.48);
\definecolor[named]{drawColor}{rgb}{0.00,0.00,0.00}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (184.81,347.78) rectangle (315.17,355.73);
\end{scope}
\end{tikzpicture}