### 2017-05-03 A5-theorem.tex

\subsection{Proof of Theorem \ref{res:oostest}}
Coverage of the confidence intervals is immediate from
Theorem~\ref{res-oost}, so we will present a proof of~\eqref{eq:23}.
By Lemma~\ref{res-convergence}, we know that $\E_R \oosA - \E_T \oosB
\to^p 0$, so
\begin{equation*}
  \Pr[\E_R \oosA \leq 0] - \Pr[\E_T \oosB \leq 0] \to 0.
\end{equation*}
Now we can use Lemma~\ref{lem-basic-coupling} to define $D_t^*$ as in
the proofs of Lemma~\ref{res-mixingale} and~\ref{res-convergence} so
that $\E_R D_t^* = \E_{t-l} D_t^*$ a.s. and
\[
\lVert D_t - D_t^* \rVert_2 \leq \couplingBound{l}
\]
and define $\oosA^* = P^{-1} \sum_{t=R+1}^{T-\h} D_t^*$. Then
\begin{equation*}
  \lVert \sqrt{P} \oosA - \sqrt{P} \oosA^* \rVert_2
  \leq P^{-1/2} \sum_{t=R+1}^{T-\h} \lVert \oosA - \oosA^* \rVert_2
  = o(1),
\end{equation*}
and, consequently, we have convergence in probability of the following
vectors:
\begin{equation*}
  (\sqrt{P} \oosA^*/\sigma,\ \E_R \oosA) -
  (\sqrt{P} \oosA/\sh,\ \E_T \oosB) \to^p 0,
\end{equation*}
where we are implicitly using consistency of $\sh^2$ for $\sigma^2$.
More importantly, this implies convergence in distribution of these
vectors, so
\begin{equation*}
  \Pr[\sqrt{P} \oosA^* / \sigma > z_\alpha
  \text{ and } \E_R \oosA \leq 0]
  - \Pr[\sqrt{P} \oosA / \sh > z_\alpha \text{ and }
  \E_T \oosB \leq 0] \to 0.
\end{equation*}

For large enough $T$, both $\Pr[\E_R \oosA \leq 0]$ and $\Pr[\E_T
\oosB \leq 0]$ are positive (the second by assumption, the first by
convergence to the second) so (for these $T$)
\begin{align}\label{eq:31}
  \Pr[\sqrt{P} \oosA^* / \sigma > z_\alpha
  \mid \E_R \oosA \leq 0] &=
  \frac{\Pr[\sqrt{P} \oosA^* / \sigma > z_\alpha \text{ and }
    \E_R \oosA \leq 0]}{\Pr[\E_R \oosA \leq 0]}
\intertext{and}\label{eq:32}
  \Pr[\sqrt{P} \oosA / \sh > z_\alpha
  \mid \E_T \oosB \leq 0] &=
  \frac{\Pr[\sqrt{P} \oosA / \sh > z_\alpha \text{ and }
    \E_T \oosB \leq 0]}{\Pr[\E_T \oosB \leq 0]}
\end{align}
almost surely. Since the terms on the RHS of Equations~\eqref{eq:31}
and~\eqref{eq:32} converge to the same limit in probability, we have
\begin{equation*}
  \Pr[\sqrt{P} \oosA / \sh > z_\alpha  \mid \E_T \oosB \leq 0]
  - \Pr[\sqrt{P} \oosA^* / \sigma > z_\alpha \mid \E_R \oosA \leq 0]
  \to^p 0
\end{equation*}
and it suffices to show that
\begin{equation*}
  \plim \Pr[\sqrt{P} \oosA^* / \sigma > z_\alpha \mid \E_R \oosA \leq 0]
  \leq \alpha.
\end{equation*}

To establish this inequality, we have
\begin{align*}
  \Pr[\sqrt{P} \oosA^* / \sigma > z_\alpha \mid \E_R \oosA \leq 0]
  &\leq \Pr[\sqrt{P} (\oosA^* - \E_R \oosA) / \sigma > z_\alpha
  \mid \E_R \oosA \leq 0] \\
  &\leq \E\big(
  \Pr[\sqrt{P} (\oosA^* - \E_R \oosA) / \sigma > z_\alpha \mid \Fs_R]
  \mid \E_R \oosA \leq 0\big).
\end{align*}
By construction,
\begin{equation*}
  \Pr[\sqrt{P} (\oosA^* - \E_R \oosA) / \sigma > z_\alpha \mid \Fs_R]
  \to \alpha
\end{equation*}
since generated pseudo OOS observations in each $D_t^*$ are
independent of $\Fs_R$, completing the proof.
\qed

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% TeX-command-extra-options: "-shell-escape"
%%% End: